### 2022-11-29 docs/cogsci_full_paper_template.tex
add latex template

% 
% Annual Cognitive Science Conference
% Sample LaTeX Paper -- Proceedings Format
%

% Original : Ashwin Ram (ashwin@cc.gatech.edu)       04/01/1994
% Modified : Johanna Moore (jmoore@cs.pitt.edu)      03/17/1995
% Modified : David Noelle (noelle@ucsd.edu)          03/15/1996
% Modified : Pat Langley (langley@cs.stanford.edu)   01/26/1997
% Latex2e corrections by Ramin Charles Nakisa        01/28/1997 
% Modified : Tina Eliassi-Rad (eliassi@cs.wisc.edu)  01/31/1998
% Modified : Trisha Yannuzzi (trisha@ircs.upenn.edu) 12/28/1999 (in process)
% Modified : Mary Ellen Foster (M.E.Foster@ed.ac.uk) 12/11/2000
% Modified : Ken Forbus                              01/23/2004
% Modified : Eli M. Silk (esilk@pitt.edu)            05/24/2005
% Modified : Niels Taatgen (taatgen@cmu.edu)         10/24/2006
% Modified : David Noelle (dnoelle@ucmerced.edu)     11/19/2014
% Modified : Roger Levy (rplevy@mit.edu)     12/31/2018

%% Change "letterpaper" in the following line to "a4paper" if you must.

\documentclass[10pt,letterpaper]{article}

\usepackage{cogsci}

\cogscifinalcopy % Uncomment this line for the final submission

\usepackage{pslatex}
\usepackage{apacite}
\usepackage{float} % Roger Levy added this and changed figure/table
                   % placement to [H] for conformity to Word template,
                   % though floating tables and figures to top is
                   % still generally recommended!

%\usepackage[none]{hyphenat} % Sometimes it can be useful to turn off
%hyphenation for purposes such as spell checking of the resulting
%PDF.  Uncomment this block to turn off hyphenation.

%\setlength\titlebox{4.5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 4.5cm (the original size).
%%If you do, we reserve the right to require you to change it back in
%%the camera-ready version, which could interfere with the timely
%%appearance of your paper in the Proceedings.

\title{Errorless irrationality: removing prediction errors from the inverse base-rate effect}

\author{{\large \bf Lenard Dome (lenarddome@gmail.com)} \\
  Brain Research and Imaging Centre \\
  University of Plymouth, Research Way, Plymouth, PL6 8BU
  \AND {\large \bf Andy J. Wills (andy.wills@plymouth.ac.uk)} \\
  Brain Research and Imaging Centre \\
  University of Plymouth, Research Way, Plymouth, PL6 8BU}

\begin{document}

\maketitle

\begin{abstract}
Include no author information in the initial submission, to facilitate
blind review.  The abstract should be one paragraph, indented 1/8~inch on both sides,
in 9~point font with single spacing. The heading ``{\bf Abstract}''
should be 10~point, bold, centered, with one line of space below
it. This one-paragraph abstract section is required only for standard
six page proceedings papers. Following the abstract should be a blank
line, followed by the header ``{\bf Keywords:}'' and a list of
descriptive keywords separated by semicolons, all in 9~point font, as
shown below.

\textbf{Keywords:} 
irrationality; prediction error; inverse base-rate effect; categorization; contingency learning
\end{abstract}

\section{Introduction}

- inverse base-rate effect
- investigating some strong assumptions about theories of the effect
- Johansen conceptual replication
- Removing error overall

\section{Experiment 1}

\subsection{Method}
\subsection{Results and Discussion}

\section{Experiment 2}

\subsection{Method}
\subsection{Results and Discussion}

\section{Discussion}

- auxiliary phenomenon (Wills CIRP addition)
- Things that any theory of IBRE should explain
- current theories fall short

\section{Open Science}

\section{Acknowledgments}

In the \textbf{initial submission}, please \textbf{do not include
  acknowledgements}, to preserve anonymity.  In the \textbf{final submission},
place acknowledgments (including funding information) in a section \textbf{at
the end of the paper}.

\bibliographystyle{apacite}

\setlength{\bibleftmargin}{.125in}
\setlength{\bibindent}{-\bibleftmargin}

\bibliography{CogSci_Template}

\end{document}

### 2022-12-19 docs/manuscript.tex
work introduction and add johansen as related work

% 
% Annual Cognitive Science Conference
% Sample LaTeX Paper -- Proceedings Format
%

% Original : Ashwin Ram (ashwin@cc.gatech.edu)       04/01/1994
% Modified : Johanna Moore (jmoore@cs.pitt.edu)      03/17/1995
% Modified : David Noelle (noelle@ucsd.edu)          03/15/1996
% Modified : Pat Langley (langley@cs.stanford.edu)   01/26/1997
% Latex2e corrections by Ramin Charles Nakisa        01/28/1997 
% Modified : Tina Eliassi-Rad (eliassi@cs.wisc.edu)  01/31/1998
% Modified : Trisha Yannuzzi (trisha@ircs.upenn.edu) 12/28/1999 (in process)
% Modified : Mary Ellen Foster (M.E.Foster@ed.ac.uk) 12/11/2000
% Modified : Ken Forbus                              01/23/2004
% Modified : Eli M. Silk (esilk@pitt.edu)            05/24/2005
% Modified : Niels Taatgen (taatgen@cmu.edu)         10/24/2006
% Modified : David Noelle (dnoelle@ucmerced.edu)     11/19/2014
% Modified : Roger Levy (rplevy@mit.edu)     12/31/2018

%% Change "letterpaper" in the following line to "a4paper" if you must.

\documentclass[10pt,letterpaper]{article}

\usepackage{cogsci}

\cogscifinalcopy % Uncomment this line for the final submission

\usepackage{pslatex}
\usepackage{apacite}
\usepackage{tikz}
\usepackage{float} % Roger Levy added this and changed figure/table
                   % placement to [H] for conformity to Word template,
                   % though floating tables and figures to top is
                   % still generally recommended!

%\usepackage[none]{hyphenat} % Sometimes it can be useful to turn off
%hyphenation for purposes such as spell checking of the resulting
%PDF.  Uncomment this block to turn off hyphenation.

%\setlength\titlebox{4.5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 4.5cm (the original size).
%%If you do, we reserve the right to require you to change it back in
%%the camera-ready version, which could interfere with the timely
%%appearance of your paper in the Proceedings.

\title{Errorless irrationality: removing prediction errors from the inverse base-rate effect}

\author{{\large \bf Lenard Dome (lenarddome@gmail.com)} \\
  Brain Research and Imaging Centre \\
  University of Plymouth, Research Way, Plymouth, PL6 8BU
  \AND {\large \bf Andy J. Wills (andy.wills@plymouth.ac.uk)} \\
  Brain Research and Imaging Centre \\
  University of Plymouth, Research Way, Plymouth, PL6 8BU}

\begin{document}

\maketitle

\begin{abstract}
Include no author information in the initial submission, to facilitate
blind review.  The abstract should be one paragraph, indented 1/8~inch on both sides,
in 9~point font with single spacing. The heading ``{\bf Abstract}''
should be 10~point, bold, centered, with one line of space below
it. This one-paragraph abstract section is required only for standard
six page proceedings papers. Following the abstract should be a blank
line, followed by the header ``{\bf Keywords:}'' and a list of
descriptive keywords separated by semicolons, all in 9~point font, as
shown below.

\textbf{Keywords:} 
irrationality; prediction error; inverse base-rate effect; categorization; contingency learning
\end{abstract}

\section{Introduction}

Inverse base-rate effect \cite<IBRE, >{medin1988problem} is an irrational tendency in humans to overweigh rare events when faced with ambiguity.
In a traditional design, people learn to categorise two overlapping sets of features under two distinct labels.
These sets share a single feature, A, and posess a unique feature, B and C, predictive of their respective category label.
During learning, these sets of features occur at different frequencies.
The features under the common label usually occur three times as much as features under the rare label \cite{kruschke1996base}.
Following training, people categorise features presented by themselves and un unique combinations.
People optimally label uniquely predictive features, B and C, presented individually with teir respective common and rare label.
Responses on the shared feature A also show base-rate following.
But when uniquely predictive features are paired, B and C, people tend to respond with the rare category label.
According Classical Probability Theory, the rational response is to attribute the common label to this ambiguous compound, because it is the most frequently occuring label.
This rare bias on ambiguous combinations of BC have been observed across different variety of different experiments and manipulations \cite{kalish2001inverse,don2017effects,don2017effects,inkster2022effect,wills2014attention}.
For a more thourough introduction into this irrational bias, see a review by \citeA{don2021attention}.

\subsection*{Assumptions of models of the IBRE}

The most prominent theories of the inverse base-rate effect involve an attentional mechanism that drives not only learning but responding as well.
These models are EXIT \cite{kruschke2001toward}, a three-layer neural network with competitive attentional gating and a four-layer neural network with an additional rapid attentional shift \cite{paskewitz2020dissecting}.
All these explanations rely on a process that reallocates attention in response to prediction errors.
Their explanation is simple.
During learning, people learn to label the AB compound first.
They are still learning to label the AC compound, so when they make an error, attention relocates towards the uniquely predictive feature C to reduce future errors.
This results in C acquiring higher attentional salience than B.
When the ambiguous BC compound is presented, C will dominate responding, resulting in an irrational tendency to respond with the rare label.
According to these models, this irrationality results from an optimisation process that tries to reduce the errors people make.

\subsection*{Current Study}

In this work, we intend to test this basic assumption of models of the IBRE.
In the following two experiments, we will gradually remove components from the design traditionally associated with prediction error.
In our first attempt, building on the observational learning condition of Experiment 2 in \citeA{johansen2007paradoxical}, we implement the canonical IBRE design with a caveat that category labels are presented in unison with features.

In our second attempt, we further remove the causal relationship between features and category labels.
The goal was to remove any design component that might affect attentional allocation.
Any presumption of causal relationship can inadvertantly relacote attention in line with the direction of causality between features and labels.

\subsection*{Related Work}

We will build on work carried out by \citeA{johansen2007paradoxical}.
\citeA{johansen2007paradoxical} failed to observe the inverse base-rate effect in an observational learning condition, where category labels were presented with features at the same time to participants.
They largely interpreted their results in opposition to EXIT \cite{kruschke2001toward}, but their experimental procedure does not correspond to the standard inverse base-rate effect procedure.
Their design involves disjoint-cues (where categories shared no features in common), while the canonical design depends on a shared features during training that facilitates attentional relocation, which in turn moves responding towards the irrational rare preference.
The only instance when they observed a rare bias was when common features presented in compound during training were paired with a rare feature presented by itself during training.
A simple explanation for that result is that the rare feature developed stronger connections with the category label.
Compound features share the connection with the category label, so any update to these connections dissipate between them.
There is no need to relocate attention to reduce errors, therefore attention will not push responding towards the rare label.

\section{Experiment 1}

Below, we detail our first attempt to test whether we could observe the rare response bias without any explicit error-driven psychological mechanism.

\subsection{Method}

\begin{table}[!ht]
  \begin{center}
    \caption{Abstract design of Experiment 1 including both test and training phases. \\}
    \label{tab:abstract-exp1}
    \begin{tabular}{llr} % text alignments
      \textbf{Training (Relative Frequencies)} & \textbf{Test}& \\
      \hline
      & \\
      $AB \to common_{1}$ (x 3) &  A, B, C,         &  \\
      $AC \to rare_{1}$   (x 1) &  AB, AC, BC      & x 20 \\
      \hline
    \end{tabular}
  \end{center}
\end{table}

\subsection{Results and Discussion}

\section{Experiment 2}

\subsection{Method}

\begin{table}[!ht]
  \begin{center}
    \caption{Abstract design of Experiment 2 including both test and training phases. X and Y are in place of the category labels. During the test phase, participants needed to select either X or Y to complete the features shown below.\\}
    \label{tab:abstract-exp2}
    \begin{tabular}{llr} % text alignments
      \textbf{Training (Relative Frequencies)} & \textbf{Test}& \\
      \hline
      & \\
      ABX x 3 &  A, B, C,         &  \\
      ACY x 1 &  AB, AC, BC      & x 20 \\
      \hline
    \end{tabular}
  \end{center}
\end{table}

\subsection{Results and Discussion}

\section{Discussion}

- auxiliary phenomenon (Wills CIRP addition)
- Things that any theory of IBRE should explain
- current theories fall short

\section{Open Science}

\section{Acknowledgments}

In the \textbf{initial submission}, please \textbf{do not include
  acknowledgements}, to preserve anonymity.  In the \textbf{final submission},
place acknowledgments (including funding information) in a section \textbf{at
the end of the paper}.

\bibliographystyle{apacite}

\setlength{\bibleftmargin}{.125in}
\setlength{\bibindent}{-\bibleftmargin}

\bibliography{library}

\end{document}

### 2022-12-19 docs/manuscript.tex
continue related work

@@
 There is no need to relocate attention to reduce errors, therefore attention will not push responding towards the rare label.
+Their design is therefore not sufficient to answer our question, whether prediction error is a necessary condition.
 
@@
 
-Below, we detail our first attempt to test whether we could observe the rare response bias without any explicit error-driven psychological mechanism.
+Below, we detail our first attempt to test whether we could observe the rare response bias without an explicit error-driven psychological mechanism.
+The design component which is most likely to result in any error-driven tuning is feedback.
+To remove feedback, we will present category labels with their respective features.
 
@@
 - current theories fall short
+- eye-tracking and attention
 

### 2023-01-03 docs/manuscript.tex
expand related work and add results

@@
 %%appearance of your paper in the Proceedings.
+\def\arraystretch{1.15}%  1 is the default, change whatever you need
 
-\title{Errorless irrationality: removing prediction errors from the inverse base-rate effect}
+\title{Errorless irrationality: removing error-driven components from the inverse base-rate effect paradigm}
 
@@
 
-We will build on work carried out by \citeA{johansen2007paradoxical}.
-\citeA{johansen2007paradoxical} failed to observe the inverse base-rate effect in an observational learning condition, where category labels were presented with features at the same time to participants.
-They largely interpreted their results in opposition to EXIT \cite{kruschke2001toward}, but their experimental procedure does not correspond to the standard inverse base-rate effect procedure.
-Their design involves disjoint-cues (where categories shared no features in common), while the canonical design depends on a shared features during training that facilitates attentional relocation, which in turn moves responding towards the irrational rare preference.
-The only instance when they observed a rare bias was when common features presented in compound during training were paired with a rare feature presented by itself during training.
-A simple explanation for that result is that the rare feature developed stronger connections with the category label.
+To our knowledge, there is only one attempt to implement the IBRE procedure without explicit feedback.
+\citeA{johansen2007paradoxical} tried to observe the inverse base-rate effect both in a predictive-learning condition and in an observational learning condition, where category labels were presented with features at the same time to participants.
+Their design involves disjoint-cues (where categories shared no features in common), while the canonical design depends on a shared feature during training that facilitates attentional relocation.
+This attentional tuning in turn pushes responding towards the rare label.
+Their design was optimised to investigate the hypothesised assymetric cognitive representation of the two categories - one of the assumptions.
+As a result, the only instance when they observed a rare bias was when common features presented in compound during training were paired with a rare feature presented by itself during training.
+This provided evidence for the assymetric cognitive represetation hypothesised to develop during training.
+In order to investigate the role of prediction error in response to feedback, we need to tweak their observational learning condition to conform to a more canonical implementation of the procedure.
+A simple neural-network explanation for the rare bias \citeA{johansen2007paradoxical} observed is that the rare feature developed stronger connections with the category label.
 Compound features share the connection with the category label, so any update to these connections dissipate between them.
-There is no need to relocate attention to reduce errors, therefore attention will not push responding towards the rare label.
-Their design is therefore not sufficient to answer our question, whether prediction error is a necessary condition.
+There is no need to relocate attention to reduce errors, therefore attention will not bias responding towards the rare label.
+
+The only two studies directly looking at error-driven processes in the IBRE are \citeA{inkster2022neural} and \citeA{wills2014attention}.
+\citeA{inkster2022neural} carried out a direct investigation into brain regions underlying error-driven learning in the IBRE.
+They observed
 
@@
       \hline
-      & \\
+      % & \\
       $AB \to common_{1}$ (x 3) &  A, B, C,         &  \\
@@
 \subsection{Results and Discussion}
+
+\begin{table}[H]
+  \begin{center}
+    \caption{Stuff. \\}
+    \label{tab:results-exp1}
+    \vskip 0.12in
+    \begin{tabular}{rcc}
+      \hline
+      & $P(common)$ & $P(rare)$ \\
+      \hline
+      A & 0.69 & 0.31\\
+      AB & 0.93 & 0.07\\
+      AC & 0.10 & 0.90\\
+      B & 0.93 & 0.07\\
+      \textbf{BC} & \textbf{0.33} & \textbf{0.66}\\
+      C & 0.06 & 0.94\\
+    \end{tabular}
+  \end{center}
+\end{table}
+
+$M = 0.66$, 95\% HDI $[0.62, 0.71]$, $\mathrm{BF}_{10} = 6.63 \times 10^{7}$
 
@@
       \hline
-      & \\
+      % & \\
       ABX x 3 &  A, B, C,         &  \\
@@
 \subsection{Results and Discussion}
+
+\begin{table}[H]
+  \begin{center}
+    \caption{Caption.\\}
+    \label{tab:results-exp2}
+    \vskip 0.12in
+    \begin{tabular}{rcc}
+      \hline
+       & $P(common)$ & $P(rare)$ \\
+      \hline
+      A & 0.76 & 0.24\\
+      AB & 0.93 & 0.07\\
+      AC & 0.11 & 0.89\\
+      B & 0.92 & 0.08\\
+      \textbf{BC} & \textbf{0.34} & \textbf{0.65}\\
+      C & 0.09 & 0.91\\
+    \end{tabular}
+  \end{center}
+\end{table}
+
+$M = 0.65$, 95\% HDI $[0.55, 0.76]$, $\mathrm{BF}_{10} = 8.13$
 

### 2023-01-03 docs/manuscript.tex
wrap up introduction and move on to experiment 1

@@
 In the following two experiments, we will gradually remove components from the design traditionally associated with prediction error.
+Our overarching goal is to investigate whether we can still observe the IBRE, even if we experimentally remove a crucial assumption of already existing accounts.
 In our first attempt, building on the observational learning condition of Experiment 2 in \citeA{johansen2007paradoxical}, we implement the canonical IBRE design with a caveat that category labels are presented in unison with features.
@@
 \citeA{johansen2007paradoxical} tried to observe the inverse base-rate effect both in a predictive-learning condition and in an observational learning condition, where category labels were presented with features at the same time to participants.
-Their design involves disjoint-cues (where categories shared no features in common), while the canonical design depends on a shared feature during training that facilitates attentional relocation.
+Their design involved disjoint-cues (where categories shared no features in common), while the canonical design depends on a shared feature during training that facilitates attentional relocation.
 This attentional tuning in turn pushes responding towards the rare label.
@@
 In order to investigate the role of prediction error in response to feedback, we need to tweak their observational learning condition to conform to a more canonical implementation of the procedure.
-A simple neural-network explanation for the rare bias \citeA{johansen2007paradoxical} observed is that the rare feature developed stronger connections with the category label.
+A simple neural-network explanation can account for this by positing that the rare feature developed stronger connections with the category label.
 Compound features share the connection with the category label, so any update to these connections dissipate between them.
 There is no need to relocate attention to reduce errors, therefore attention will not bias responding towards the rare label.
+But stronger $C \to rare$ feature-label connections will result in a rare bias on BC compunds during test.
+Our current study modified the canonical IBRE procedure in an attempt to remove error-driven components from the design - which includes a shared cue.
+This contrasts \citeA{johansen2007paradoxical}, who tried to remove the shared cue that pushed attention towards C during training and resulted in the assymetric representation.
+Here, we strictly focus on prediction error in relation to behaviour.
+% I will also need to include Experiment 3 listed shared cue condition
 
+% I need to go through some studies because I think there are at least one more, but I 
 The only two studies directly looking at error-driven processes in the IBRE are \citeA{inkster2022neural} and \citeA{wills2014attention}.
+\citeA{wills2014attention} observed posterior selection negativity and concurrent frontal positivity for C relative to B, which gave evidence for an error-driven selective attentional learning process.
 \citeA{inkster2022neural} carried out a direct investigation into brain regions underlying error-driven learning in the IBRE.
-They observed
+Their ROI analysis explicitly targeted areas that were hypothesised to be involved in the computation of prediction error.
+They showed that these areas exhibited greater activation during the test phase for C relative to B with a presence of a shared cue during training.
+Both \citeA{wills2014attention} and \citeA{inkster2022neural} gave strong evidence in support of error-driven attentional learning accounts of the IBRE.
+In our study, we will look for the effect while trying to take out prediction error from the experimental design.
+In contrast, they looked for the neural substrates of error-driven accounts of the effect.
 
@@
 To remove feedback, we will present category labels with their respective features.
+We retain the sequential property of the experiment, which means that participants learn about feature and category relationships on a trial by trial basis.
+% work out how models might predict rational responding
 
 \subsection{Method}
+
+\subsubsection*{Participants}
+
+Recruited 169 participants.
+
+\subsubsection*{Stimuli}
+
+\subsubsection*{Procedure}
 
@@
 
+\subsubsection*{Exclusion}
+
+To match performance with
+
+\subsubsection*{Analysis}
+
 \subsection{Results and Discussion}
+
+125 ppt made it into the analysis.
 
@@
       \hline
-      A & 0.69 & 0.31\\
-      AB & 0.93 & 0.07\\
-      AC & 0.10 & 0.90\\
-      B & 0.93 & 0.07\\
-      \textbf{BC} & \textbf{0.33} & \textbf{0.66}\\
-      C & 0.06 & 0.94\\
+      A & 0.69 & 0.31 \\
+      AB & 0.94 & 0.06 \\
+      AC & 0.08 & 0.92 \\
+      B & 0.94 & 0.06 \\
+      \textbf{BC} & \textbf{0.33} & \textbf{0.67} \\
+      C & 0.04 & 0.96 \\
     \end{tabular}
@@
 
-$M = 0.66$, 95\% HDI $[0.62, 0.71]$, $\mathrm{BF}_{10} = 6.63 \times 10^{7}$
+$M = 0.67$, 95\% HDI $[0.62, 0.72]$, $\mathrm{BF}_{10} = 1.11 \times 10^{7}$
 
@@
 \subsection{Method}
+
+\subsubsection*{Participants}
+
+Recruited 65 participants.
+
+\subsubsection*{Stimuli}
+
+\subsubsection*{Procedure}
 
@@
 
+30 made it into the analysis
+
 \begin{table}[H]
@@
       \hline
-      A & 0.76 & 0.24\\
-      AB & 0.93 & 0.07\\
-      AC & 0.11 & 0.89\\
-      B & 0.92 & 0.08\\
-      \textbf{BC} & \textbf{0.34} & \textbf{0.65}\\
-      C & 0.09 & 0.91\\
+      A & 0.78 & 0.22  \\
+      AB & 0.95 & 0.05 \\
+      AC & 0.09 & 0.91 \\
+      B & 0.92 & 0.07  \\
+      \textbf{BC} & \textbf{0.35} & \textbf{0.65} \\
+      C & 0.08 & 0.92 \\
     \end{tabular}
@@
 
-$M = 0.65$, 95\% HDI $[0.55, 0.76]$, $\mathrm{BF}_{10} = 8.13$
+$M = 0.64$, 95\% HDI $[0.53, 0.75]$, $\mathrm{BF}_{10} = 4.09$
 

### 2023-01-04 docs/manuscript.tex
work on experiment 1

@@
 \documentclass[10pt,letterpaper]{article}
+\usepackage{hyperref}
 
@@
 
-Inverse base-rate effect \cite<IBRE, >{medin1988problem} is an irrational tendency in humans to overweigh rare events when faced with ambiguity.
+The \textit{inverse base-rate effect} \cite<IBRE, >{medin1988problem} is an irrational tendency in humans to overweigh rare events when faced with ambiguity.
 In a traditional design, people learn to categorise two overlapping sets of features under two distinct labels.
@@
 
+Experiment 1 is a conceptual replication of an experiment included in the Appendix of \citeA{johansen2007paradoxical}, but not the main text.
+There are no details about the procedure of this experiment, so we cannot make direct comparisons.
+The only information available are the list of test items (23), the doubled-up design (2 sets of categories and features), and the sample size.
+We substantially simplified our implementation by removing the doubled-up design and reducing the number of test items to 6.
+
 \subsection{Method}
@@
 
-Recruited 169 participants.
+Participants were undergraduate students who recieved course credit for their participantion.
+We recruited 169 participants online through the SONA recruitment system.
+
+\subsubsection*{Apparatus}
+
+The experiment was programmed in JsPsych \cite{deleeuw2015JsPsych} to be run in a web browser.
+The experiment code is avalaible at \href{https://osf.io/auwvt/?view_only=2dc8384074fa4bcf9f2e3937fdaee2b4}{OSF} and \href{https://github.com/lenarddome/ply216-observational-ibre}{GitHub}.
+% links to be inserted here
+Participants completed the experiment on their personal computer.
+The experiment did not allow the use of tablets and smartphones.
 
@@
 
+Category labels corresponded with response keys and were called Disease \textbf{Z} and Disease \textbf{L}.
+Category features were symptoms: fever, headache, and rash.
+These physical features were randomly allocated to abstract features, A, B and C at the beginning of each individual session.
+Features and labels appeared in full sentences, such as '\textit{John has fever and rash, which belongs to disease Z}'.
+Names were randomly drawn from a pool of male and female first names.
+The list were compiled from an online repository of popular baby names\footnote[1]{The list was taken and later curated from \href{https://github.com/aruljohn/popular-baby-names}{a GitHub repository}.}.
+We selected the 50 most popular male and female names from 2021.
+
 \subsubsection*{Procedure}
+
+Table \ref{tab:abstract-exp1} summarises the abstract design of the experiment.
+This design is the simplest implementation of the IBRE procedure to date.
+Participants completed two phases: a training and test phase.
+In the training phase, they encountered descriptions of people, the symptoms they experienced, and their respective disease.
+These descriptions appeared in the format of '\textit{John has fever and rash, which belongs to disease Z}'.
+Partipants studied these examples and when they were ready to move on, they pressed the spacebar.
+They needed to complete reading the description within 5 seconds.
+If the 5 seconds threshold has passed, a screen appeared with the message '\textit{Please respond faster!}'.
+In each training block, participants encountered 6 common diseases (common category exemplars) and 2 rare diseases (rare category exemplars).
+After the second block of training, participants were given a choice.
+They could either move straight to the test phase or complete another training block.
+There were a maximum of 5 blocks they could complete.
 
@@
 
-To match performance with
+To match performance with the predictive learning implementations of the IBRE, we decided to exclude participants whose test performance on the training items were below 0.75 accuracy.
+We arrived at this threshold by testing all difference levels of accuracy through calculating the Bayes factors for binomial data.
+We used the method implemented in BayesFactor R package \cite{morey2022bayes}.
+If the Bayes Factor fell above 3, we concluded that we have sufficient evidence to believe the participant learned the training items.
 
@@
 
+In order to test the presence of the IBRE, we calculated a Bayes Factor for a one-sample design.
+We calculate the probability of responding rare on the critical BC test item, $P(rare|BC)$, for each participant.
+Then we tested this distribution of probabilities against the null, $mu = 0.5$, which denoted random responding.
+If the Bayes Factor fell below 1/3, we concluded that participants' responses are not different from random responding.
+If the Bayes Factor fell above 3, we concluded that participants' responses reliably differ from null.
+If the mean probability of $P(rare|BC)$ is higher then 0.5, we conclude that we observed the IBRE.
+Values lower than 0.5 would indicate rational responding.
+We used the method implemented in the BayesFactor R package \cite{morey2022bayes}.
+
 \subsection{Results and Discussion}
 
-125 ppt made it into the analysis.
+After exclusion, 125 participants made it into our main analysis.
 
@@
   \begin{center}
-    \caption{Stuff. \\}
+    \caption{Group-level mean probabilities for each stimuli presented during the test phase in Experiment 1. \\}
     \label{tab:results-exp1}

### 2023-01-05 docs/manuscript.tex
finish experiment 2 results and discussion

@@
 Include no author information in the initial submission, to facilitate
-blind review.  The abstract should be one paragraph, indented 1/8~inch on both sides,
-in 9~point font with single spacing. The heading ``{\bf Abstract}''
-should be 10~point, bold, centered, with one line of space below
-it. This one-paragraph abstract section is required only for standard
-six page proceedings papers. Following the abstract should be a blank
-line, followed by the header ``{\bf Keywords:}'' and a list of
-descriptive keywords separated by semicolons, all in 9~point font, as
-shown below.
+blind review.
 
@@
 In a traditional design, people learn to categorise two overlapping sets of features under two distinct labels.
-These sets share a single feature, A, and posess a unique feature, B and C, predictive of their respective category label.
+These sets share a single feature, A, and possess a unique feature, B and C, predictive of their respective category label.
 During learning, these sets of features occur at different frequencies.
@@
 Following training, people categorise features presented by themselves and un unique combinations.
-People optimally label uniquely predictive features, B and C, presented individually with teir respective common and rare label.
+People optimally label uniquely predictive features, B and C, presented individually with their respective common and rare labels.
 Responses on the shared feature A also show base-rate following.
 But when uniquely predictive features are paired, B and C, people tend to respond with the rare category label.
-According Classical Probability Theory, the rational response is to attribute the common label to this ambiguous compound, because it is the most frequently occuring label.
-This rare bias on ambiguous combinations of BC have been observed across different variety of different experiments and manipulations \cite{kalish2001inverse,don2017effects,don2017effects,inkster2022effect,wills2014attention}.
-For a more thourough introduction into this irrational bias, see a review by \citeA{don2021attention}.
+According to Classical Probability Theory, the rational response is to attribute the common label to this ambiguous compound, because it is the most frequently occuring label.
+This rare bias on ambiguous combinations of BC has been observed across a different variety experimental manipulations \cite{kalish2001inverse,don2017effects,don2017effects,inkster2022effect,wills2014attention}.
+For a more thorough introduction to this irrational bias, see a review by \citeA{don2021attention}.
 
@@
 According to these models, this irrationality results from an optimisation process that tries to reduce the errors people make.
+This process results in an asymmetric representation that can be summarised as AB belongs to common, $AB \to common$, and C belongs to rare, $C \to rare$.
 
@@
 In our second attempt, we further remove the causal relationship between features and category labels.
-The goal was to remove any design component that might affect attentional allocation.
-Any presumption of causal relationship can inadvertantly relacote attention in line with the direction of causality between features and labels.
+The goal was to remove any design component that might affect attentional allocation or development of assymmetric representation in response to errors.
+Any presumption of a causal relationship can inadvertently relocate attention in line with the direction of causality between features and labels.
 
@@
 To our knowledge, there is only one attempt to implement the IBRE procedure without explicit feedback.
-\citeA{johansen2007paradoxical} tried to observe the inverse base-rate effect both in a predictive-learning condition and in an observational learning condition, where category labels were presented with features at the same time to participants.
-Their design involved disjoint-cues (where categories shared no features in common), while the canonical design depends on a shared feature during training that facilitates attentional relocation.
-This attentional tuning in turn pushes responding towards the rare label.
-Their design was optimised to investigate the hypothesised assymetric cognitive representation of the two categories - one of the assumptions.
+In their Experiment 2, \citeA{johansen2007paradoxical} tried to observe the inverse base-rate effect both in a predictive-learning condition and in an observational learning condition, where category labels were presented with features at the same time to participants.
+Their design involved disjoint cues (where categories shared no features in common), while the canonical design depends on a shared feature during training that facilitates attentional relocation.
+This attentional tuning in turn pushes responding toward the rare label.
+Their design was optimized to investigate the hypothesised asymmetric cognitive representation of the two categories.
 As a result, the only instance when they observed a rare bias was when common features presented in compound during training were paired with a rare feature presented by itself during training.
-This provided evidence for the assymetric cognitive represetation hypothesised to develop during training.
+This provided evidence for the asymmetric cognitive representation hypothesised to develop during training.
+A simple neural-network explanation can account for this by positing that the rare feature developed stronger connections with the category label.
+Compound features share the connection with the category label, so any update to these connections dissipates between them.
+There is no need to relocate attention to reduce errors, therefore attention will not bias responding toward the rare label.
+But stronger $C \to rare$ feature-label connections will result in a rare bias on BC compounds during the test phase.
 In order to investigate the role of prediction error in response to feedback, we need to tweak their observational learning condition to conform to a more canonical implementation of the procedure.
-A simple neural-network explanation can account for this by positing that the rare feature developed stronger connections with the category label.
-Compound features share the connection with the category label, so any update to these connections dissipate between them.
-There is no need to relocate attention to reduce errors, therefore attention will not bias responding towards the rare label.
-But stronger $C \to rare$ feature-label connections will result in a rare bias on BC compunds during test.
+\citeA{johansen2007paradoxical} included the result of a short pilot experiment in their Appendix.
+There are no details about the procedure of this experiment, so we cannot make direct comparisons.
+We will follow up on that line of investigation.
+
 Our current study modified the canonical IBRE procedure in an attempt to remove error-driven components from the design - which includes a shared cue.
-This contrasts \citeA{johansen2007paradoxical}, who tried to remove the shared cue that pushed attention towards C during training and resulted in the assymetric representation.
-Here, we strictly focus on prediction error in relation to behaviour.
+This contrasts Experiment 2 in \citeA{johansen2007paradoxical}, who tried to remove the shared cue that pushed attention towards C during training and resulted in the assymetric representation.
+Here, we strictly focus on prediction error concerning behaviour.
 % I will also need to include Experiment 3 listed shared cue condition
 
-% I need to go through some studies because I think there are at least one more, but I 
+% I need to go through some studies because I think there is at least one more
 The only two studies directly looking at error-driven processes in the IBRE are \citeA{inkster2022neural} and \citeA{wills2014attention}.
@@
 To remove feedback, we will present category labels with their respective features.
-We retain the sequential property of the experiment, which means that participants learn about feature and category relationships on a trial by trial basis.
+We retain the sequential property of the experiment, which means that participants learn about feature and category relationships on a trial-by-trial basis.
 % work out how models might predict rational responding
 
-Experiment 1 is a conceptual replication of an experiment included in the Appendix of \citeA{johansen2007paradoxical}, but not the main text.
-There are no details about the procedure of this experiment, so we cannot make direct comparisons.
-The only information available are the list of test items (23), the doubled-up design (2 sets of categories and features), and the sample size.
+Experiment 1 is a conceptual replication of an experiment included in the Appendix of \citeA{johansen2007paradoxical}.
+The only information available is the list of test items (23), the doubled-up design (2 sets of categories and features), and the sample size.
 We substantially simplified our implementation by removing the doubled-up design and reducing the number of test items to 6.
@@
 
-Participants were undergraduate students who recieved course credit for their participantion.
+Participants were undergraduate students who received course credit for their participation.
 We recruited 169 participants online through the SONA recruitment system.
@@
 The experiment code is avalaible at \href{https://osf.io/auwvt/?view_only=2dc8384074fa4bcf9f2e3937fdaee2b4}{OSF} and \href{https://github.com/lenarddome/ply216-observational-ibre}{GitHub}.
-% links to be inserted here
-Participants completed the experiment on their personal computer.
+Participants completed the experiment on their personal computers.
 The experiment did not allow the use of tablets and smartphones.
@@
 Category features were symptoms: fever, headache, and rash.
-These physical features were randomly allocated to abstract features, A, B and C at the beginning of each individual session.
+These physical features were randomly allocated to abstract features, A, B, and C at the beginning of each session.
 Features and labels appeared in full sentences, such as '\textit{John has fever and rash, which belongs to disease Z}'.
 Names were randomly drawn from a pool of male and female first names.
-The list were compiled from an online repository of popular baby names\footnote[1]{The list was taken and later curated from \href{https://github.com/aruljohn/popular-baby-names}{a GitHub repository}.}.
+The list was compiled from an online repository of popular baby names\footnote[1]{The list was taken and later curated from \href{https://github.com/aruljohn/popular-baby-names}{a GitHub repository}.}.
 We selected the 50 most popular male and female names from 2021.
+Disease names corresponded to response keys and were randomly allocated to either the common or rare category label at the beginning of each session.
 
@@
 This design is the simplest implementation of the IBRE procedure to date.
-Participants completed two phases: a training and test phase.
-In the training phase, they encountered descriptions of people, the symptoms they experienced, and their respective disease.
+Participants completed two phases: a training and a test phase.
+In the training phase, they encountered descriptions of people, the symptoms they experienced, and their respective diseases.
 These descriptions appeared in the format of '\textit{John has fever and rash, which belongs to disease Z}'.
-Partipants studied these examples and when they were ready to move on, they pressed the spacebar.
+Participants studied these examples and when they were ready to move on, they pressed the spacebar.
 They needed to complete reading the description within 5 seconds.
@@
 There were a maximum of 5 blocks they could complete.
+
+In the test phase, participants judged individual symptoms and novel combinations of old symptoms, see Table \ref*{tab:abstract-exp1}.
+Symptoms appeared in a sentence, such as '\textit{John has a fever.}', with a prompt asking participants to label what disease the person has, '\textit{Does the patient has disease Z or disease L?}'.
+Participants had to respond by pressing either Z or L on the keyboard.
+They had 10 seconds to do so, otherwise, a '\textit{Please respond faster!}' message appeared.
+After the button press, there was no feedback,
+Each unique test item and training item (occurring in the test phase) was repeated 20 times.
 
@@
 
-\subsubsection*{Exclusion}
-
-To match performance with the predictive learning implementations of the IBRE, we decided to exclude participants whose test performance on the training items were below 0.75 accuracy.
-We arrived at this threshold by testing all difference levels of accuracy through calculating the Bayes factors for binomial data.
-We used the method implemented in BayesFactor R package \cite{morey2022bayes}.
-If the Bayes Factor fell above 3, we concluded that we have sufficient evidence to believe the participant learned the training items.
-
 \subsubsection*{Analysis}
@@
 In order to test the presence of the IBRE, we calculated a Bayes Factor for a one-sample design.
-We calculate the probability of responding rare on the critical BC test item, $P(rare|BC)$, for each participant.
+We calculate the probability of responding with the rare label on the critical BC test item, $P(rare|BC)$, for each participant.
 Then we tested this distribution of probabilities against the null, $mu = 0.5$, which denoted random responding.
@@
 If the Bayes Factor fell above 3, we concluded that participants' responses reliably differ from null.
-If the mean probability of $P(rare|BC)$ is higher then 0.5, we conclude that we observed the IBRE.
+If the mean probability of $P(rare|BC)$ is higher than 0.5, we conclude that we observed the IBRE.
 Values lower than 0.5 would indicate rational responding.
 We used the method implemented in the BayesFactor R package \cite{morey2022bayes}.
+
+\subsubsection*{Exclusion}
+
+To match performance with the predictive learning implementations of the IBRE, we decided to exclude participants whose test performance on the training items fell below 0.75 accuracy.
+We arrived at this threshold by testing all different levels of accuracy by calculating the Bayes factors for binomial data.
+We used the method implemented in BayesFactor R package \cite{morey2022bayes}.
+If the Bayes Factor fell above 3, we concluded that we have sufficient evidence to believe the participant learned the training items.
 
@@
 After exclusion, 125 participants made it into our main analysis.
+In a summary, the qualitative pattern in our results corresponds to the base result of the IBRE.
+Table \ref*{tab:results-exp1} shows the group-level probabilities for each item.
+Predictive features and training items are classified into their respective category.
+Participants exhibited a reliable common preference for $A$, $M_{A} = 0.68$, 95\% HDI $[0.63, 0.73]$, $\mathrm{BF}_{10} = 2.45 \times 10^{7}$.
+People explicitly followed the base rate  - responded rationally according to Probability Theory.
+On the contrary, participants showed a reliable rare preference for $BC$, $M_{BC} = 0.67$, 95\% HDI $[0.62, 0.72]$, $\mathrm{BF}_{10} = 1.11 \times 10^{7}$.
+This gives us a sufficient amount of evidence to conclude that we have observed the IBRE.
 
-\begin{table}[H]
+\begin{table}[ht]
   \begin{center}
-    \caption{Group-level mean probabilities for each stimuli presented during the test phase in Experiment 1. \\}
+    \caption{Group-level mean probabilities for each stimulus presented during the test phase in Experiment 1. \\}
     \label{tab:results-exp1}
@@
 
-$M = 0.67$, 95\% HDI $[0.62, 0.72]$, $\mathrm{BF}_{10} = 1.11 \times 10^{7}$
+Here, we report a successful conceptual replication of an observational learning IBRE procedure reported by \cite{johansen2007paradoxical}.
+In the current experimental design, the IBRE emerged in the absence of an explicit prediction error that drives the development of attentional allocation.
+This prediction error also drives the development of an asymmetric cognitive representation.
+All current theories of the IBRE rely on the assumption that this irrational rare preference arises as a result of optimising accuracy during the training phase.
+In the absence of this explicit prediction error, current theories cannot deal with encoding this type of input representation.
+
+One shortcoming of the current design is that participants can still make predictions about feature--category on a trial-by-trial basis.
+Given that the general assumption is that diseases cause symptoms, participants could likely assume a causal link between symptoms and diseases.
+This assumed causal relationship can encourage participants to make not an explicit (responding with the category label via the keyboard) but an implicit prediction.
+Informally, participants might think of a certain feature--label causal relationship while reading the sentences.
+People then resolve errors between the expected and the observed feature--label causality by allocating attention to rare features to distinguish diseases.
+A simple solution is to remove any design component that makes it clear to participants what the category label is.
+In addition, stimuli need to be constructed in a way that reduces the chance of people assuming any causal relationship between its features.
 
 \section{Experiment 2}
+
+In this experiment, we implemented the IBRE in a way most similar to memory experiments.
 
@@
 \subsubsection*{Stimuli}
+
+\begin{figure}
+  \begin{center}
+    \caption{Stuff.}
+    \label{figure:exp2-stimuli}
+    \includegraphics[scale=0.15]{figures/experiment_2_stimuli.pdf}
+  \end{center}
+\end{figure}
 
@@
 
+- asymmetric representation as an assumption
+
+Asymmetric representation can arise in different ways.
+One way it is manifested is the attentional tuning of cognitive representation of category exemplars.
+Another scenario is
+
 - auxiliary phenomenon (Wills CIRP addition)
@@
 In the \textbf{initial submission}, please \textbf{do not include
-  acknowledgements}, to preserve anonymity.  In the \textbf{final submission},
+  acknowledgments}, to preserve anonymity.  In the \textbf{final submission},
 place acknowledgments (including funding information) in a section \textbf{at

### 2023-01-05 docs/manuscript.tex
start experiment 2

@@
 People then resolve errors between the expected and the observed feature--label causality by allocating attention to rare features to distinguish diseases.
+
 A simple solution is to remove any design component that makes it clear to participants what the category label is.
@@
 
-In this experiment, we implemented the IBRE in a way most similar to memory experiments.
+In this experiment, we implemented the IBRE in a way most similar to cued-recall memory tasks.
+Previous category labels were treated as features.
+And features were selected to be solid black geometric shapes.
+The task asked participants to memorise the arrangement of these shapes.
+On each trial, we randomised the position of the geometric shapes in the arrangement.
+This further minimised the chances of having any design component suggesting which feature is the category label.
 
@@
 
-Recruited 65 participants.
+We recruited 65 undergraduate students who completed the exeriment for partial course credit.
+Recruitment was done via the SONA recruitment system.
 

### 2023-01-09 docs/manuscript.tex
work on experiment 2 method and results

@@
   \begin{center}
-    \caption{Group-level mean probabilities for each stimulus presented during the test phase in Experiment 1. \\}
+    \caption{Group-level mean probabilities for each stimulus presented during the test phase in Experiment 1 after exclusion. \\}
     \label{tab:results-exp1}
@@
 On each trial, we randomised the position of the geometric shapes in the arrangement.
-This further minimised the chances of having any design component suggesting which feature is the category label.
+This further minimised the chances of having any design component suggestive of which feature is the category label.
 
@@
   \begin{center}
-    \caption{Stuff.}
+    \caption{Simple geometric shapes used as stimuli in Experiment 2.}
     \label{figure:exp2-stimuli}
@@
 
+I have a myriads of things to say here. \\
+
+\subsubsection*{Analysis and Exclusion}
+
+We applied the same analysis and exclusion methods as in Experiment 1.
+
 \subsection{Results and Discussion}
 
-30 made it into the analysis
+After exclusion, 30 participants made it into our analysis.
+The group-level mean probabilities are shown on Table \ref{tab:results-exp2}.
+The resuls are a qualitative and ordinal match to Experiment 1.
+Participants showed a clear common preference for stimuli A, $M_{A} = 0.21$, 95\% HDI $[0.11, 0.31]$, $\mathrm{BF}_{10} = 11,312.94$.
 
@@
   \begin{center}
-    \caption{Caption.\\}
+    \caption{Group-level mean probabilities for each stimulus
+    presented during the test phase in Experiment 2 after exclusion.\\}
     \label{tab:results-exp2}
@@
 
-$M = 0.64$, 95\% HDI $[0.53, 0.75]$, $\mathrm{BF}_{10} = 4.09$
+Participants also showed a reliable rare preference on ambiguous BC trials, $M_{BC} = 0.64$, 95\% HDI $[0.53, 0.75]$, $\mathrm{BF}_{10} = 4.09$
+Therefore we conclude that we reliably detected the IBRE.
 
@@
 
-\section{Acknowledgments}
-
-In the \textbf{initial submission}, please \textbf{do not include
-  acknowledgments}, to preserve anonymity.  In the \textbf{final submission},
-place acknowledgments (including funding information) in a section \textbf{at
-the end of the paper}.
+All things will be free.
 

### 2023-01-09 docs/manuscript.tex
finish experiment 2 and start discussion

@@
 \documentclass[10pt,letterpaper]{article}
+
 \usepackage{hyperref}
@@
 But when uniquely predictive features are paired, B and C, people tend to respond with the rare category label.
-According to Classical Probability Theory, the rational response is to attribute the common label to this ambiguous compound, because it is the most frequently occuring label.
-This rare bias on ambiguous combinations of BC has been observed across a different variety experimental manipulations \cite{kalish2001inverse,don2017effects,don2017effects,inkster2022effect,wills2014attention}.
-For a more thorough introduction to this irrational bias, see a review by \citeA{don2021attention}.
+According to Classical Probability Theory, the rational response is to attribute the common label to this ambiguous compound, because it is the most frequently occurring label.
+This rare bias on ambiguous combinations of BC has been observed across a different variety of experimental manipulations \cite{kalish2001inverse,don2017effects,don2021attention,inkster2022effect,wills2014attention}.
+For a more thorough introduction to this irrational bias, see a review by \citeA{don2021hearing}.
 
-\subsection*{Assumptions of models of the IBRE}
+\subsection{Assumptions of theories of the IBRE}
 
 The most prominent theories of the inverse base-rate effect involve an attentional mechanism that drives not only learning but responding as well.
-These models are EXIT \cite{kruschke2001toward}, a three-layer neural network with competitive attentional gating and a four-layer neural network with an additional rapid attentional shift \cite{paskewitz2020dissecting}.
-All these explanations rely on a process that reallocates attention in response to prediction errors.
+These theories are formal models.
+They are: a neural network with an exemplar-mediated attention to distinctive input, EXIT \cite{kruschke2001toward}, a three-layer neural network with competitive attentional gating, and a four-layer neural network with an additional rapid attentional shift \cite{paskewitz2020dissecting}.
+All these explanations rely on a process that reallocates attention in response to prediction errors - they update attentional values according to gradient descent.
 Their explanation is simple.
@@
 According to these models, this irrationality results from an optimisation process that tries to reduce the errors people make.
-This process results in an asymmetric representation that can be summarised as AB belongs to common, $AB \to common$, and C belongs to rare, $C \to rare$.
+This process results in an asymmetric representation that can be summarised as AB belongs to common, $AB \to common$, and C belongs to rare, $C \to rare$ \cite{kruschke2001inverse}.
 
-\subsection*{Current Study}
+\subsection{Current Study}
 
-In this work, we intend to test this basic assumption of models of the IBRE.
+In this work, we intend to test this basic assumption of these theories.
 In the following two experiments, we will gradually remove components from the design traditionally associated with prediction error.
@@
 In our second attempt, we further remove the causal relationship between features and category labels.
-The goal was to remove any design component that might affect attentional allocation or development of assymmetric representation in response to errors.
+The goal was to remove any design component that might affect attentional allocation or the development of asymmetric representation in response to errors.
 Any presumption of a causal relationship can inadvertently relocate attention in line with the direction of causality between features and labels.
 
-\subsection*{Related Work}
+\subsection{Related Work}
 
 To our knowledge, there is only one attempt to implement the IBRE procedure without explicit feedback.
-In their Experiment 2, \citeA{johansen2007paradoxical} tried to observe the inverse base-rate effect both in a predictive-learning condition and in an observational learning condition, where category labels were presented with features at the same time to participants.
-Their design involved disjoint cues (where categories shared no features in common), while the canonical design depends on a shared feature during training that facilitates attentional relocation.
+In their Experiment 2, \citeA{johansen2007paradoxical} tried to observe the inverse base-rate effect both in a predictive-learning condition and in an observational learning condition.
+In the observational learning condition, category labels and features were presented at the same time to participants.
+Their design involved disjoint cues (where categories shared no features in common), while the canonical design depends on a shared feature during training.
+According to the theories of the IBRE, this shared feature facilitates attentional relocation.
 This attentional tuning in turn pushes responding toward the rare label.
+In contrast, \citeA{johansen2007paradoxical} trained participants on $AB \to common$ and $C \to rare$.
+% LD: maybe include here some info on the shared-cue effect?
 Their design was optimized to investigate the hypothesised asymmetric cognitive representation of the two categories.
-As a result, the only instance when they observed a rare bias was when common features presented in compound during training were paired with a rare feature presented by itself during training.
-This provided evidence for the asymmetric cognitive representation hypothesised to develop during training.
+As a result, the only instance when they observed a rare bias was when the common feature, $B$ ,presented in compound during training was paired with a rare, $C$, feature presented by itself during training.
+This provided evidence for the asymmetric cognitive representation \cite{kruschke2001inverse} hypothesised to develop during training.
+% LD: this speculation might be redundant?
 A simple neural-network explanation can account for this by positing that the rare feature developed stronger connections with the category label.
@@
 In order to investigate the role of prediction error in response to feedback, we need to tweak their observational learning condition to conform to a more canonical implementation of the procedure.
-\citeA{johansen2007paradoxical} included the result of a short pilot experiment in their Appendix.
+This will involve reimplementing the shared feature.
+In terms of a clear observational-learning version of the IBRE, \citeA{johansen2007paradoxical} included the result of a short pilot experiment in their Appendix.
 There are no details about the procedure of this experiment, so we cannot make direct comparisons.
@@
 Our current study modified the canonical IBRE procedure in an attempt to remove error-driven components from the design - which includes a shared cue.
-This contrasts Experiment 2 in \citeA{johansen2007paradoxical}, who tried to remove the shared cue that pushed attention towards C during training and resulted in the assymetric representation.
+This contrasts Experiment 2 in \citeA{johansen2007paradoxical}, who tried to remove the shared cue that pushed attention towards C during training and resulted in the asymetric representation.
 Here, we strictly focus on prediction error concerning behaviour.
-% I will also need to include Experiment 3 listed shared cue condition
+% LD: I will also need to include Experiment 3 listed shared cue condition?
 
-% I need to go through some studies because I think there is at least one more
+% LD: I need to go through some studies because I think there is at least one more
 The only two studies directly looking at error-driven processes in the IBRE are \citeA{inkster2022neural} and \citeA{wills2014attention}.
@@
 We retain the sequential property of the experiment, which means that participants learn about feature and category relationships on a trial-by-trial basis.
-% work out how models might predict rational responding
+% LD: work out how models might predict rational responding?
 
@@
 
-\subsubsection*{Participants}
+\subsubsection{Participants}
 
@@
 
-\subsubsection*{Apparatus}
+\subsubsection{Apparatus}
 
 The experiment was programmed in JsPsych \cite{deleeuw2015JsPsych} to be run in a web browser.
-The experiment code is avalaible at \href{https://osf.io/auwvt/?view_only=2dc8384074fa4bcf9f2e3937fdaee2b4}{OSF} and \href{https://github.com/lenarddome/ply216-observational-ibre}{GitHub}.
 Participants completed the experiment on their personal computers.
@@
 
-\subsubsection*{Stimuli}
+\subsubsection{Stimuli}
 
@@
 
-\subsubsection*{Procedure}
+\subsubsection{Procedure}
 
@@
 Each unique test item and training item (occurring in the test phase) was repeated 20 times.
+The test phase, therefore, included 120 trials, which were broken down into 5 blocks of 24 trials.
 
@@
 
-\subsubsection*{Analysis}
+\subsubsection{Analysis}
 
@@
 
-\subsubsection*{Exclusion}
+\subsubsection{Exclusion}
 
@@
 
-In this experiment, we implemented the IBRE in a way most similar to cued-recall memory tasks.
+In this experiment, we implemented the IBRE in a way most similar to cued recall tasks.
 Previous category labels were treated as features.
@@
 
-\subsubsection*{Participants}
+\subsubsection{Participants}
 
-We recruited 65 undergraduate students who completed the exeriment for partial course credit.
-Recruitment was done via the SONA recruitment system.
+We recruited 65 undergraduate students who completed the experiment for partial course credit.
+Recruitment was done via the SONA recruitment system. \\
 
-\subsubsection*{Stimuli}
+\subsubsection{Stimuli}
 
@@
 
-\subsubsection*{Procedure}
+Stimuli were common solid geometric shapes, shown in Figure \ref*{figure:exp2-stimuli}.
+Common and rare category labels were turned into features X and Y respectively.
+Each shape was randomly allocated to one of the abstract features shown in Table \ref*{tab:abstract-exp2}.
+
+\subsubsection{Procedure}
 
@@
 
-I have a myriads of things to say here. \\
+Table \ref*{tab:abstract-exp2} depicts the abstract experiment design.
+Similar to the previous experiment, participants completed two phases: an encoding/training and a cued-recall/test phase.
+In the training/encoding phase, participants were repeatedly exposed to the exemplars and were asked to memorise the arrangement of geometric shapes.
+Compared to Experiment 1, exemplars were composed of three geometric shapes.
+On each trial, geometric shapes appeared in random order so the position of features on the screen was completely counterbalanced.
+This resulted in 24 trials within each block, which contained 18 common trials and 6 rare trials.
+Similar to Experiment 1, participants could complete a maximum of 5 blocks.
+After the first block, they were given a chance after completing each block to move straight to the test phase.
+The trial structure and response deadlines corresponded to Experiment 1.
 
-\subsubsection*{Analysis and Exclusion}
+In the test phase, participants were shown \textit{incomplete} arrangement of geometric shapes and were asked to complete them.
+On each test trial, they were asked to select either \textbf{X} or \textbf{Y} to complete the arrangement.
+Similar to Experiment 1, each test item (incomplete arrangement of shapes) appeared 20 times. 
+Similarly, the test phase was composed of 120 trials presented across 5 blocks of 24 trials.
+
+\subsubsection{Analysis and Exclusion}
 
@@
 After exclusion, 30 participants made it into our analysis.
-The group-level mean probabilities are shown on Table \ref{tab:results-exp2}.
-The resuls are a qualitative and ordinal match to Experiment 1.
-Participants showed a clear common preference for stimuli A, $M_{A} = 0.21$, 95\% HDI $[0.11, 0.31]$, $\mathrm{BF}_{10} = 11,312.94$.
+The group-level mean probabilities are shown in Table \ref{tab:results-exp2}.
+The results are a qualitative and ordinal match to Experiment 1.
+Participants showed a clear common preference for stimuli A, $M_{A} = 0.77$, 95\% HDI $[0.68, 0.87]$, $\mathrm{BF}_{10} = 10,316.41$.
 
@@
   \begin{center}
-    \caption{Group-level mean probabilities for each stimulus
-    presented during the test phase in Experiment 2 after exclusion.\\}
+    \caption{Group-level mean probabilities for each stimulus presented during the test phase in Experiment 2 after exclusion.\\}
     \label{tab:results-exp2}
@@
 Participants also showed a reliable rare preference on ambiguous BC trials, $M_{BC} = 0.64$, 95\% HDI $[0.53, 0.75]$, $\mathrm{BF}_{10} = 4.09$
-Therefore we conclude that we reliably detected the IBRE.
+This gives us a sufficient amount of evidence to conclude that we have observed the IBRE.
+
+Here, we further demonstrated that the IBRE can arise without experimental-design components that explicitly promote an error-driven process.
+In addition, Experiment 2 suggests that the IBRE generalises beyond simple predictive-learning \cite{kruschke1996base,don2021hearing} and decision-making \cite{johansen2007paradoxical} paradigms.
 
@@
 
+In this study, we tested a central assumption of the most prominent theories of the IBRE.
+This central assumption was prediction error.
+In our first attempt, we implemented an observational learning version of the canonical IBRE procedure.
+This meant that features and category labels appeared on the screen at the same time.
+Participants learned about categories by reading complete sentences that describe what symptoms people exhibit and what diseases they have.
+The experiment included no feedback and required no responses from participants during training.
+From a theoretical perspective, there was no opportunity for making an explicit error, therefore the condition for prediction error.
+Regardless, we observed the inverse base-rate effect.
+One limitation of this approach was that there are assumed causal relationships between features (symptoms) and labels (diseases).
+These relationships can predispose participants to make feature-to-label predictions, which will results in prediction error and attentional relocation.
+In our second attempt, we further removed the causal relationship between features and labels by changing the stimuli and their presentation.
+Here, participants saw nothing but an arrangement of geometric shapes, where previous category labels were treated as features.
+There were no causal links between features and labels.
+When participants needed to complete incomplete arrangements of these shapes, they still exhibited a rare bias on $BC$ trials.
+So, we still observed the IBRE.
+
+% LD: so it seems that johansen et al treated asymmetric representation as an experimental condition and not as a theoretical construct
+% LD: I think this is a problem, because asymmetric representation is hypothesized to be a asymmetry in the way the two categories represented
 - asymmetric representation as an assumption
 
-Asymmetric representation can arise in different ways.
-One way it is manifested is the attentional tuning of cognitive representation of category exemplars.
-Another scenario is
+% Asymmetric representation can arise in different ways.
+The hypothesised way asymmetric representation is manifested is the attentional tuning of cognitive representation of category exemplars.
 
@@
 - eye-tracking and attention
+- attention can still be allocated asymmetrically to drive the distinguishability of categories.
+
+\section{Conclusion}
 
@@
 
-All things will be free.
+We made available the two experiments written in javascript, the analysis code, the raw data, and all other supplementary materials.
+Repositories related to Experiment 1 can be found on \href{https://osf.io/auwvt/?view_only=2dc8384074fa4bcf9f2e3937fdaee2b4}{OSF}, and \href{https://github.com/lenarddome/ply216-observational-ibre}{GitHub}.
+Repositories related to Experiment 2 can be found on \href{https://osf.io/2tmc4/?view_only=489ebc888ef84a6b9b904072cfbe74df}{OSF} and \href{https://github.com/lenarddome/ply222-non-causal-ibre}{Github}.
+The main repository that includes this manuscript and links to the materials for the two experiments can be found on \href{https://github.com/lenarddome/pu093-errorless-attentionless-ibre}{GitHub}.
 

### 2023-01-10 docs/manuscript.tex
continue working on discussion

@@
 
-\cogscifinalcopy % Uncomment this line for the final submission
+% \cogscifinalcopy % Uncomment this line for the final submission
 
@@
 When participants needed to complete incomplete arrangements of these shapes, they still exhibited a rare bias on $BC$ trials.
-So, we still observed the IBRE.
+We still observed the IBRE.
 
-% LD: so it seems that johansen et al treated asymmetric representation as an experimental condition and not as a theoretical construct
-% LD: I think this is a problem, because asymmetric representation is hypothesized to be a asymmetry in the way the two categories represented
-- asymmetric representation as an assumption
+The two experiments suggest that the necessary conditions to observe the IBRE are fewer than previously established.
+In Experiment 2, the only remaining conditions are the two uniquely predictive features, an overlapping feature, sequential presentation and the base rate.
+\cite{johansen2007paradoxical} gave evidence that even the overlapping features are not necessary.
+In their design, common category exemplars were made up of two features, while rare exemplars only had a single feature.
+This can mean that common exemplars need to be composed of at least two features, while the rare exemplar must have only one feature.
+If the rare exemplar has one uniquely predictive feature and another feature that is not shared with the other category, we do not observe the effect.
+This is the shared-cue effect \cite{kruschke2001inverse, wills2014attention}.
+One can hypothesise that a shared cue is only a sufficient condition.
+In addition, one can hypothesise that in the absence of a shared feature, common exemplars must contain more features than rare exemplars.
+
+% LD: so it seems that Johansen et al treated asymmetric representation as an experimental condition and not as a theoretical construct
+% LD: I think this is a problem because asymmetric representation is hypothesized to be an asymmetry in the way the two categories represented
+Compared to \citeA{johansen2007paradoxical}, we do not consider asymmetric representation and base-rate neglect as necessary conditions.
+From our point of view, they result from the experimental conditions, which was our focus here.
 
@@
 We made available the two experiments written in javascript, the analysis code, the raw data, and all other supplementary materials.
-Repositories related to Experiment 1 can be found on \href{https://osf.io/auwvt/?view_only=2dc8384074fa4bcf9f2e3937fdaee2b4}{OSF}, and \href{https://github.com/lenarddome/ply216-observational-ibre}{GitHub}.
-Repositories related to Experiment 2 can be found on \href{https://osf.io/2tmc4/?view_only=489ebc888ef84a6b9b904072cfbe74df}{OSF} and \href{https://github.com/lenarddome/ply222-non-causal-ibre}{Github}.
+Experiment 1 is shared via \href{https://osf.io/auwvt/?view_only=2dc8384074fa4bcf9f2e3937fdaee2b4}{OSF}, and \href{https://github.com/lenarddome/ply216-observational-ibre}{GitHub}.
+Experiment 2 is shared via \href{https://osf.io/2tmc4/?view_only=489ebc888ef84a6b9b904072cfbe74df}{OSF} and \href{https://github.com/lenarddome/ply222-non-causal-ibre}{Github}.
 The main repository that includes this manuscript and links to the materials for the two experiments can be found on \href{https://github.com/lenarddome/pu093-errorless-attentionless-ibre}{GitHub}.

### 2023-01-10 docs/manuscript.tex
transitional commit

@@
 Here, we further demonstrated that the IBRE can arise without experimental-design components that explicitly promote an error-driven process.
-In addition, Experiment 2 suggests that the IBRE generalises beyond simple predictive-learning \cite{kruschke1996base,don2021hearing} and decision-making \cite{johansen2007paradoxical} paradigms.
 
@@
 One limitation of this approach was that there are assumed causal relationships between features (symptoms) and labels (diseases).
-These relationships can predispose participants to make feature-to-label predictions, which will results in prediction error and attentional relocation.
+These relationships can predispose participants to make feature-to-label predictions, which will result in prediction error and attentional relocation.
 In our second attempt, we further removed the causal relationship between features and labels by changing the stimuli and their presentation.
@@
 \cite{johansen2007paradoxical} gave evidence that even the overlapping features are not necessary.
-In their design, common category exemplars were made up of two features, while rare exemplars only had a single feature.
+In their design, common category exemplars were made up of two features, while rare exemplars only had a single feature - disjoint cues.
 This can mean that common exemplars need to be composed of at least two features, while the rare exemplar must have only one feature.
@@
 This is the shared-cue effect \cite{kruschke2001inverse, wills2014attention}.
-One can hypothesise that a shared cue is only a sufficient condition.
+One can hypothesise that the shared cue and the disjoint cue are sufficient conditions, but at least one of them is necessary.
 In addition, one can hypothesise that in the absence of a shared feature, common exemplars must contain more features than rare exemplars.
@@
 % LD: I think this is a problem because asymmetric representation is hypothesized to be an asymmetry in the way the two categories represented
-Compared to \citeA{johansen2007paradoxical}, we do not consider asymmetric representation and base-rate neglect as necessary conditions.
-From our point of view, they result from the experimental conditions, which was our focus here.
+Compared to \citeA{johansen2007paradoxical}, we do not consider the asymmetric representation and the base-rate neglect as necessary conditions.
+From our point of view, they are theoretical constructs.
+They are not part of the experimental design.
+We consider them to result from the precise environmental conditions required for the IBRE as opposed to being environmental conditions necessary/sufficient for the IBRE.
 
@@
 \section{Conclusion}
+
+In addition, Experiment 2 suggests that the IBRE generalises beyond simple predictive-learning \cite{kruschke1996base,don2021hearing} and decision-making \cite{johansen2007paradoxical} paradigms.
 

### 2023-01-10 docs/manuscript.tex
broaden discussion

@@
 This central assumption was prediction error.
+
 In our first attempt, we implemented an observational learning version of the canonical IBRE procedure.
@@
 These relationships can predispose participants to make feature-to-label predictions, which will result in prediction error and attentional relocation.
+
 In our second attempt, we further removed the causal relationship between features and labels by changing the stimuli and their presentation.
@@
 We still observed the IBRE.
+
+In both experiments, the IBRE occured without any explicit detail in the experimental procedure that would result in prediction error.
+Therefore, any theorised error-driven process must be able to operate without explicit feedback.
+Most prominent theories and their corresponding formal specification rely on relocating attention in response to prediction error.
+They are unable to process the current experiment, because they are not designed to encode information presented without feedback or fit cued recall tasks.
 
@@
 They are not part of the experimental design.
-We consider them to result from the precise environmental conditions required for the IBRE as opposed to being environmental conditions necessary/sufficient for the IBRE.
+We consider them to result from the environmental conditions specified in the experiment required for the IBRE as opposed to being necessary/sufficient conditions for the IBRE.
 
@@
 
-In addition, Experiment 2 suggests that the IBRE generalises beyond simple predictive-learning \cite{kruschke1996base,don2021hearing} and decision-making \cite{johansen2007paradoxical} paradigms.
+In Experiment 1, we conducted a successful conceputal replication of \citeA{johansen2007paradoxical}, which gave evidence for the IBRE being independent of supervised learning procedures.
+In addition, Experiment 2 further suggests that the IBRE generalises beyond simple predictive-learning \cite{kruschke1996base,don2021hearing} and decision-making \cite{johansen2007paradoxical} paradigms.
+Theories of IBRE are inadequate to account for these findings, because of their inability to extend beyond supervised learning.
 

### 2023-01-10 docs/manuscript.tex
include future research

@@
 % Asymmetric representation can arise in different ways.
-The hypothesised way asymmetric representation is manifested is the attentional tuning of cognitive representation of category exemplars.
+One hypothesised way asymmetric representation is manifested is the attentional tuning of cognitive representation of category exemplars.
+This is not necessarily absent in our experiment, but is not directly tested.
+Our experiments do not give direct evidence against the role of attention in developing the asymmetric representation.
+Nonetheless, it must not happen through an error-driven process as conceptualised in most prominent theories.
+To further investigate this, the cued-recall IBRE procedure could incorporate eye-tracking to measure dwell time and order of information encoding.
+Even if we observe more and longer fixations to C relative to B, attention can still be allocated asymmetrically to drive the distinguishability of categories.
+Explanations do not need to invoke an explicit error-driven process.
 
@@
 - eye-tracking and attention
-- attention can still be allocated asymmetrically to drive the distinguishability of categories.
+-
 

### 2023-01-11 docs/manuscript.tex
write abstract and do small edits

@@
 \begin{abstract}
-Include no author information in the initial submission, to facilitate
-blind review.
+
+The inverse base-rate effect is a robust irrational bias that arises when people face ambiguity.
+The most prominent theories of this irrational bias depend on prediction error.
+In this study, we gradually removed elements of a predictive learning design.
+In our first experiment, we removed explicit feedback by implementing the inverse base-rate effect in an observational learning procedure. 
+In our second study, we further removed any causal relationship between stimulus features and category labels.
+This removed any information participants could use to identify category labels.
+In both experiments, the inverse base-rate effect persisted and remained robust.
+This outcome suggests that this irrational bias is independent of supervised learning procedures.
+We propose that any theories and models of the inverse base-rate effect must be designed in a way that allows information encoding and connection updates to proceed without prediction error.
+We end by proposing two clear paths for future investigations.
 
@@
 The features under the common label usually occur three times as much as features under the rare label \cite{kruschke1996base}.
-Following training, people categorise features presented by themselves and un unique combinations.
-People optimally label uniquely predictive features, B and C, presented individually with their respective common and rare labels.
-Responses on the shared feature A also show base-rate following.
-But when uniquely predictive features are paired, B and C, people tend to respond with the rare category label.
+Following training, people categorise features presented by themselves and in novel combinations.
+People tend to optimally label uniquely predictive features, $B$ and $C$, with their respective common and rare labels when presented by themselves.
+Responses on the shared feature A tend to show base-rate following.
+But when uniquely predictive features are paired, $BC$, people tend to respond with the rare category label.
 According to Classical Probability Theory, the rational response is to attribute the common label to this ambiguous compound, because it is the most frequently occurring label.
@@
 
+% LD: if I talk about supervised/predictive learning at the end, I should probably include something here
 The most prominent theories of the inverse base-rate effect involve an attentional mechanism that drives not only learning but responding as well.
@@
 They are: a neural network with an exemplar-mediated attention to distinctive input, EXIT \cite{kruschke2001toward}, a three-layer neural network with competitive attentional gating, and a four-layer neural network with an additional rapid attentional shift \cite{paskewitz2020dissecting}.
-All these explanations rely on a process that reallocates attention in response to prediction errors - they update attentional values according to gradient descent.
+All these explanations rely on a process that relocates attention in response to prediction errors - they update attentional values according to gradient descent.
 Their explanation is simple.
@@
 In their Experiment 2, \citeA{johansen2007paradoxical} tried to observe the inverse base-rate effect both in a predictive-learning condition and in an observational learning condition.
-In the observational learning condition, category labels and features were presented at the same time to participants.
-Their design involved disjoint cues (where categories shared no features in common), while the canonical design depends on a shared feature during training.
+In the observational learning condition, category labels and features appeared together at the same time on the computer screen.
+Their design involved disjoint cues (where categories shared no features in common), while the canonical design depends on a shared feature during training \cite{kruschke2001inverse, wills2014attention}.
 According to the theories of the IBRE, this shared feature facilitates attentional relocation.
@@
 
-In both experiments, the IBRE occured without any explicit detail in the experimental procedure that would result in prediction error.
+In both experiments, the IBRE occurred without any explicit detail in the experimental procedure that would result in prediction error.
 Therefore, any theorised error-driven process must be able to operate without explicit feedback.
 Most prominent theories and their corresponding formal specification rely on relocating attention in response to prediction error.
-They are unable to process the current experiment, because they are not designed to encode information presented without feedback or fit cued recall tasks.
+They are unable to process the current experiment because they are not designed to encode information presented without feedback or fit cued recall tasks.
 
 The two experiments suggest that the necessary conditions to observe the IBRE are fewer than previously established.
+% LD: so it seems that Johansen et al treated asymmetric representation as an experimental condition and not as a theoretical construct
+% LD: I think this is a problem because asymmetric representation is hypothesized to be an asymmetry in the way the two categories represented
+Compared to \citeA{johansen2007paradoxical}, we also do not consider asymmetric representation and base-rate neglect as necessary conditions.
+We consider them to result from the environmental conditions specified in the experiment required for the IBRE as opposed to being necessary/sufficient conditions for the IBRE.
+From our point of view, they are theoretical constructs.
+They are not part of the experimental design.
+
 In Experiment 2, the only remaining conditions are the two uniquely predictive features, an overlapping feature, sequential presentation and the base rate.
@@
 
-% LD: so it seems that Johansen et al treated asymmetric representation as an experimental condition and not as a theoretical construct
-% LD: I think this is a problem because asymmetric representation is hypothesized to be an asymmetry in the way the two categories represented
-Compared to \citeA{johansen2007paradoxical}, we do not consider the asymmetric representation and the base-rate neglect as necessary conditions.
-From our point of view, they are theoretical constructs.
-They are not part of the experimental design.
-We consider them to result from the environmental conditions specified in the experiment required for the IBRE as opposed to being necessary/sufficient conditions for the IBRE.
-
 % Asymmetric representation can arise in different ways.
 One hypothesised way asymmetric representation is manifested is the attentional tuning of cognitive representation of category exemplars.
-This is not necessarily absent in our experiment, but is not directly tested.
-Our experiments do not give direct evidence against the role of attention in developing the asymmetric representation.
+This is not necessarily absent in our experiments but is not directly tested.
+Our experiments do not give direct evidence against the role of attention in developing asymmetric representation.
 Nonetheless, it must not happen through an error-driven process as conceptualised in most prominent theories.
@@
 
-- auxiliary phenomenon (Wills CIRP addition)
-- Things that any theory of IBRE should explain
-- current theories fall short
-- eye-tracking and attention
--
-
 \section{Conclusion}
 
+Across two experiments, we investigated whether the IBRE requires error-driven components of the experimental procedure.
 In Experiment 1, we conducted a successful conceputal replication of \citeA{johansen2007paradoxical}, which gave evidence for the IBRE being independent of supervised learning procedures.
-In addition, Experiment 2 further suggests that the IBRE generalises beyond simple predictive-learning \cite{kruschke1996base,don2021hearing} and decision-making \cite{johansen2007paradoxical} paradigms.
+In addition, Experiment 2 further suggests that the IBRE generalises beyond simple predictive-learning \cite<e.g.
+>{medin1988problem,kruschke1996base,wills2014attention} and decision-making \cite{johansen2007paradoxical} paradigms.
+This further suggest prediction error in terms of explicit feedback is not a necessary condition.
 Theories of IBRE are inadequate to account for these findings, because of their inability to extend beyond supervised learning.

### 2023-01-11 docs/manuscript.tex
transitional commit
correcting phrasing

@@
 The most prominent theories of this irrational bias depend on prediction error.
-In this study, we gradually removed elements of a predictive learning design.
+In this study, we gradually removed elements of a predictive learning design to test the extent to which error-driven processes underly this bias.
 In our first experiment, we removed explicit feedback by implementing the inverse base-rate effect in an observational learning procedure. 
-In our second study, we further removed any causal relationship between stimulus features and category labels.
+In our second study, we further removed any causal relationship between stimulus features and category labels by moving towards an unsupervised learning procedure.
 This removed any information participants could use to identify category labels.
@@
 This outcome suggests that this irrational bias is independent of supervised learning procedures.
-We propose that any theories and models of the inverse base-rate effect must be designed in a way that allows information encoding and connection updates to proceed without prediction error.
+We propose that any theories and models of the inverse base-rate effect must manage information encoding and connection updates without explicit prediction error.
 We end by proposing two clear paths for future investigations.
@@
 The \textit{inverse base-rate effect} \cite<IBRE, >{medin1988problem} is an irrational tendency in humans to overweigh rare events when faced with ambiguity.
-In a traditional design, people learn to categorise two overlapping sets of features under two distinct labels.
+In a traditional design, people learn to categorise two overlapping sets of features under two distinct category labels.
 These sets share a single feature, A, and possess a unique feature, B and C, predictive of their respective category label.
@@
 People tend to optimally label uniquely predictive features, $B$ and $C$, with their respective common and rare labels when presented by themselves.
-Responses on the shared feature A tend to show base-rate following.
+Responses on the shared feature $A$ tend to show base-rate following.
 But when uniquely predictive features are paired, $BC$, people tend to respond with the rare category label.
-According to Classical Probability Theory, the rational response is to attribute the common label to this ambiguous compound, because it is the most frequently occurring label.
+According to Classical Probability Theory, the rational response is to categorise this ambiguous combination under the common label, because it is the most frequently occurring label.
 This rare bias on ambiguous combinations of BC has been observed across a different variety of experimental manipulations \cite{kalish2001inverse,don2017effects,don2021attention,inkster2022effect,wills2014attention}.
@@
 According to these models, this irrationality results from an optimisation process that tries to reduce the errors people make.
-This process results in an asymmetric representation that can be summarised as AB belongs to common, $AB \to common$, and C belongs to rare, $C \to rare$ \cite{kruschke2001inverse}.
+This process creates an asymmetric cognitive representation that can be summarised as AB belongs to common, $AB \to common$, and C belongs to rare, $C \to rare$ \cite{kruschke2001inverse}.
 
@@
 To our knowledge, there is only one attempt to implement the IBRE procedure without explicit feedback.
-In their Experiment 2, \citeA{johansen2007paradoxical} tried to observe the inverse base-rate effect both in a predictive-learning condition and in an observational learning condition.
+In terms of a clear observational-learning version of the IBRE, \citeA{johansen2007paradoxical} included the result of a short pilot experiment in their Appendix.
+There are no details about the procedure of this experiment, so we cannot make direct comparisons.
+Nonetheless, in their Experiment 2, \citeA{johansen2007paradoxical} tried to observe the inverse base-rate effect both in a predictive-learning condition and in an observational-learning condition.
 In the observational learning condition, category labels and features appeared together at the same time on the computer screen.
-Their design involved disjoint cues (where categories shared no features in common), while the canonical design depends on a shared feature during training \cite{kruschke2001inverse, wills2014attention}.
+Their design involved disjoint cues where categories shared no features in common.
+The canonical IBRE design involves a shared feature during training \cite{kruschke2001inverse, wills2014attention}.
 According to the theories of the IBRE, this shared feature facilitates attentional relocation.
@@
 In contrast, \citeA{johansen2007paradoxical} trained participants on $AB \to common$ and $C \to rare$.
+Similarly, the $common$ category occured three times as often as $rare$.
 % LD: maybe include here some info on the shared-cue effect?
@@
 % LD: this speculation might be redundant?
-A simple neural-network explanation can account for this by positing that the rare feature developed stronger connections with the category label.
+
+A simple neural-network explanation can account for \citeA{johansen2007paradoxical}'s.
+In a simple network, the rare feature develops stronger connections with the category label.
 Compound features share the connection with the category label, so any update to these connections dissipates between them.
@@
 But stronger $C \to rare$ feature-label connections will result in a rare bias on BC compounds during the test phase.
+This is strong evidence for the assymetric cognitive representation, but it still builds on the presence of explicit prediction error.
 In order to investigate the role of prediction error in response to feedback, we need to tweak their observational learning condition to conform to a more canonical implementation of the procedure.
 This will involve reimplementing the shared feature.
-In terms of a clear observational-learning version of the IBRE, \citeA{johansen2007paradoxical} included the result of a short pilot experiment in their Appendix.
-There are no details about the procedure of this experiment, so we cannot make direct comparisons.
-We will follow up on that line of investigation.
-
-Our current study modified the canonical IBRE procedure in an attempt to remove error-driven components from the design - which includes a shared cue.
-This contrasts Experiment 2 in \citeA{johansen2007paradoxical}, who tried to remove the shared cue that pushed attention towards C during training and resulted in the asymetric representation.
-Here, we strictly focus on prediction error concerning behaviour.
+Our current study modified the canonical IBRE procedure in an attempt to remove components from the design that is associated with explicit prediction error.
+This contrasts Experiment 2 in \citeA{johansen2007paradoxical}, who tried to remove the shared cue that pushed attention towards C during training in response to prediction error.
+Here, we strictly focus on prediction error as the element of the experimental procedure.
 % LD: I will also need to include Experiment 3 listed shared cue condition?
@@
 Both \citeA{wills2014attention} and \citeA{inkster2022neural} gave strong evidence in support of error-driven attentional learning accounts of the IBRE.
-In our study, we will look for the effect while trying to take out prediction error from the experimental design.
-In contrast, they looked for the neural substrates of error-driven accounts of the effect.
+In our study, we will look for the effect while trying reduce the chance that cognitive processes calculate prediction error through modifying the experimental design.
+In contrast, they looked for the neural substrates of error-driven processes incoporated into explanations of the effect.
+% This presumes that prediction error is necessary for the inverse base-rate effect.
+% Arguably, one can predict that if there is no clear explicit design component (feedback) promoting prediction error as conceptualised in the models, the IBRE will not occur.
 
@@
 Each unique test item and training item (occurring in the test phase) was repeated 20 times.
-The test phase, therefore, included 120 trials, which were broken down into 5 blocks of 24 trials.
+So, the test phase included 120 trials, which were broken down into 5 blocks of 24 trials.
 
@@
 
-In this experiment, we implemented the IBRE in a way most similar to cued recall tasks.
+In this experiment, we implemented the IBRE in a way most similar to cued-recall tasks.
 Previous category labels were treated as features.
@@
 Table \ref*{tab:abstract-exp2} depicts the abstract experiment design.
-Similar to the previous experiment, participants completed two phases: an encoding/training and a cued-recall/test phase.
+Similar to the previous experiment, participants completed two phases: an encoding/training and a test phase.
 In the training/encoding phase, participants were repeatedly exposed to the exemplars and were asked to memorise the arrangement of geometric shapes.
@@
 Most prominent theories and their corresponding formal specification rely on relocating attention in response to prediction error.
-They are unable to process the current experiment because they are not designed to encode information presented without feedback or fit cued recall tasks.
+They are unable to process the current experiment because they are not designed to encode information presented without feedback.
 
@@
 Nonetheless, it must not happen through an error-driven process as conceptualised in most prominent theories.
-To further investigate this, the cued-recall IBRE procedure could incorporate eye-tracking to measure dwell time and order of information encoding.
+To further investigate this, the cued-recall procedure could incorporate eye-tracking to measure dwell time and order of information encoding.
 Even if we observe more and longer fixations to C relative to B, attention can still be allocated asymmetrically to drive the distinguishability of categories.

### 2023-01-13 docs/manuscript.tex
abstract and removed cued recall references
(i am unclear ATM what was my reasoning behind calling it cued recall)

--- 2023-01-11 docs/manuscript.tex
@@
 The most prominent theories of this irrational bias depend on prediction error.
-In this study, we gradually removed elements of a predictive learning design.
+In this study, we gradually detached elements of a predictive learning design to test the extent to which error-driven processes underly this bias.
 In our first experiment, we removed explicit feedback by implementing the inverse base-rate effect in an observational learning procedure. 
-In our second study, we further removed any causal relationship between stimulus features and category labels.
+In our second study, we further removed any causal relationship between stimulus features and category labels by moving towards an unsupervised learning procedure.
 This removed any information participants could use to identify category labels.
@@
 This contrasts Experiment 2 in \citeA{johansen2007paradoxical}, who tried to remove the shared cue that pushed attention towards C during training and resulted in the asymetric representation.
-Here, we strictly focus on prediction error concerning behaviour.
+Here, we strictly focus on prediction error as the element of the experimental procedure.
 % LD: I will also need to include Experiment 3 listed shared cue condition?
@@
 
-In this experiment, we implemented the IBRE in a way most similar to cued recall tasks.
+In this experiment, we implemented the IBRE in a way most similar to unsupervised category learning paradigms.
 Previous category labels were treated as features.
@@
 Nonetheless, it must not happen through an error-driven process as conceptualised in most prominent theories.
-To further investigate this, the cued-recall IBRE procedure could incorporate eye-tracking to measure dwell time and order of information encoding.
+To further investigate this, the unsupervised IBRE procedure could incorporate eye-tracking to measure dwell time and order of information encoding.
 Even if we observe more and longer fixations to C relative to B, attention can still be allocated asymmetrically to drive the distinguishability of categories.

### 2023-01-13 docs/manuscript.tex

--- 2023-01-11 docs/manuscript.tex
@@
 The most prominent theories of this irrational bias depend on prediction error.
-In this study, we gradually removed elements of a predictive learning design to test the extent to which error-driven processes underly this bias.
+In this study, we gradually detached elements of a predictive learning design to test the extent to which error-driven processes underly this bias.
 In our first experiment, we removed explicit feedback by implementing the inverse base-rate effect in an observational learning procedure. 
@@
 Each unique test item and training item (occurring in the test phase) was repeated 20 times.
-So, the test phase included 120 trials, which were broken down into 5 blocks of 24 trials.
+The test phase, therefore, included 120 trials, which were broken down into 5 blocks of 24 trials.
 
@@
 
-In this experiment, we implemented the IBRE in a way most similar to cued-recall tasks.
+In this experiment, we implemented the IBRE in a way most similar to unsupervised category learning paradigms.
 Previous category labels were treated as features.
@@
 Nonetheless, it must not happen through an error-driven process as conceptualised in most prominent theories.
-To further investigate this, the cued-recall procedure could incorporate eye-tracking to measure dwell time and order of information encoding.
+To further investigate this, the unsupervised IBRE procedure could incorporate eye-tracking to measure dwell time and order of information encoding.
 Even if we observe more and longer fixations to C relative to B, attention can still be allocated asymmetrically to drive the distinguishability of categories.

### 2023-01-13 docs/manuscript.tex
reverting some changes
responding to Andy's review

--- 2023-01-11 docs/manuscript.tex
@@
 In a traditional design, people learn to categorise two overlapping sets of features under two distinct category labels.
-These sets share a single feature, A, and possess a unique feature, B and C, predictive of their respective category label.
+These sets share a single feature, $A$, and possess a unique feature, $B$ and $C$, predictive of their respective category label.
+The training thus can be summarised under two trial types, which we will express as $AB \to common$ and $AC \to rare$.
 During learning, these sets of features occur at different frequencies.
@@
 Their explanation is simple.
-During learning, people learn to label the AB compound first.
-They are still learning to label the AC compound, so when they make an error, attention relocates towards the uniquely predictive feature C to reduce future errors.
-This results in C acquiring higher attentional salience than B.
-When the ambiguous BC compound is presented, C will dominate responding, resulting in an irrational tendency to respond with the rare label.
+During learning, people learn to label the $AB$ compound first, but they are still learning to label the $AC$ compound.
+The presence of $A$ tend to push participants to generalise what they learned about $AB$, so they label $AC$ as common, which results in an error.
+After making this error, attention relocates towards the uniquely predictive feature $C$ to reduce future errors.
+This results in $C$ acquiring higher attentional salience than $B$.
+When the ambiguous $BC$ compound is presented, this attentional allocation presists and thus $C$ will dominate responding.
+This results in an irrational tendency to respond with the rare label.
 According to these models, this irrationality results from an optimisation process that tries to reduce the errors people make.
-This process creates an asymmetric cognitive representation that can be summarised as AB belongs to common, $AB \to common$, and C belongs to rare, $C \to rare$ \cite{kruschke2001inverse}.
+This process creates an asymmetric cognitive representation that can be summarised as $AB$ belongs to common, $AB \to common$, and C belongs to rare, $C \to rare$ \cite{kruschke2001inverse}.
 
@@
 The goal was to remove any design component that might affect attentional allocation or the development of asymmetric representation in response to errors.
-Any presumption of a causal relationship can inadvertently relocate attention in line with the direction of causality between features and labels.
+Any presumption of a causal relationship might inadvertently relocate attention in line with the direction of causality between features and labels.
 

### 2023-01-13 docs/manuscript.tex
complete reviews on page 2, 3 and 4

@@
 The design component which is most likely to result in any error-driven tuning is feedback.
-To remove feedback, we will present category labels with their respective features.
+To remove feedback, Experiment 1 will present category labels simultaneously with their respective features.
 We retain the sequential property of the experiment, which means that participants learn about feature and category relationships on a trial-by-trial basis.
@@
 Names were randomly drawn from a pool of male and female first names.
-The list was compiled from an online repository of popular baby names\footnote[1]{The list was taken and later curated from \href{https://github.com/aruljohn/popular-baby-names}{a GitHub repository}.}.
+The list was compiled from an online repository of popular baby names\footnote[1]{The list was taken and later curated from a GitHub repository: \href{https://github.com/aruljohn/popular-baby-names}{https://github.com/aruljohn/popular-baby-names}.}.
 We selected the 50 most popular male and female names from 2021.
@@
 They could either move straight to the test phase or complete another training block.
+A prompt appeared saying that '\textit{Now you have the option to skip the rest of the training phase and move straight to the test phase.
+If you think you need some more time, you can continue training and study more patients.}'.
 There were a maximum of 5 blocks they could complete.
@@
 In the test phase, participants judged individual symptoms and novel combinations of old symptoms, see Table \ref*{tab:abstract-exp1}.
-Symptoms appeared in a sentence, such as '\textit{John has a fever.}', with a prompt asking participants to label what disease the person has, '\textit{Does the patient has disease Z or disease L?}'.
+Symptoms appeared in a sentence, such as '\textit{John has a fever.}', with a prompt asking participants to label what disease the person has, '\textit{Does the patient have disease Z or disease L?}'.
 Participants had to respond by pressing either Z or L on the keyboard.
@@
 To match performance with the predictive learning implementations of the IBRE, we decided to exclude participants whose test performance on the training items fell below 0.75 accuracy.
-We arrived at this threshold by testing all different levels of accuracy by calculating the Bayes factors for binomial data.
-We used the method implemented in BayesFactor R package \cite{morey2022bayes}.
-If the Bayes Factor fell above 3, we concluded that we have sufficient evidence to believe the participant learned the training items.
+This level of accuracy was the lowest at which the evidence that the participant performed better than chance was above the Bayes Factor of 3.
+We calculated the Bayes Factor for binomial proportions via the method implemented in BayesFactor R package \cite{morey2022bayes}.
 
@@
 After exclusion, 125 participants made it into our main analysis.
-In a summary, the qualitative pattern in our results corresponds to the base result of the IBRE.
+In summary, the qualitative pattern in our results corresponds to the base result of the IBRE.
 Table \ref*{tab:results-exp1} shows the group-level probabilities for each item.
@@
 Participants exhibited a reliable common preference for $A$, $M_{A} = 0.68$, 95\% HDI $[0.63, 0.73]$, $\mathrm{BF}_{10} = 2.45 \times 10^{7}$.
-People explicitly followed the base rate  - responded rationally according to Probability Theory.
-On the contrary, participants showed a reliable rare preference for $BC$, $M_{BC} = 0.67$, 95\% HDI $[0.62, 0.72]$, $\mathrm{BF}_{10} = 1.11 \times 10^{7}$.
+For this cue, people explicitly followed the base rate - responded rationally according to Probability Theory.
+In contrast, participants showed a reliable rare preference for $BC$, $M_{BC} = 0.67$, 95\% HDI $[0.62, 0.72]$, $\mathrm{BF}_{10} = 1.11 \times 10^{7}$.
 This gives us a sufficient amount of evidence to conclude that we have observed the IBRE.
@@
 
-Here, we report a successful conceptual replication of an observational learning IBRE procedure reported by \cite{johansen2007paradoxical}.
+Thus the current study strongly confirms that the IBRE can be observed in an observational procedure.
 In the current experimental design, the IBRE emerged in the absence of an explicit prediction error that drives the development of attentional allocation.
-This prediction error also drives the development of an asymmetric cognitive representation.
-All current theories of the IBRE rely on the assumption that this irrational rare preference arises as a result of optimising accuracy during the training phase.
-In the absence of this explicit prediction error, current theories cannot deal with encoding this type of input representation.
+% This prediction error also drives the development of an asymmetric cognitive representation.
+All EXIT-like theories of the IBRE rely on the assumption that this irrational rare preference arises as a result of optimising accuracy during the training phase.
+In the absence of this explicit prediction error, EXIT-like theories cannot predict the presence of the IBRE.
 
-One shortcoming of the current design is that participants can still make predictions about feature--category on a trial-by-trial basis.
+One aspect of the current design is that participants can still experience internally-generated prediction errors from feature to categories on a trial-by-trial basis.
 Given that the general assumption is that diseases cause symptoms, participants could likely assume a causal link between symptoms and diseases.
@@
 
-A simple solution is to remove any design component that makes it clear to participants what the category label is.
-In addition, stimuli need to be constructed in a way that reduces the chance of people assuming any causal relationship between its features.
+In Experiment 2, we adress this by removing any design component that makes it clear to participants what the category label is.
+And we also use stimuli that reduces the chance of people assuming any causal relationship between its features.
 
@@
 Similar to Experiment 1, each test item (incomplete arrangement of shapes) appeared 20 times.
+Various arrangement of shapes appeared in the middle of the screen.
+The response options X and Y with the corresponding shapes were shown below.
+The prompt asked participants to pick one of the shapes to complete the arrangement.
+Participants could respond by pressing either X or Y on the keyboard.
 Similarly, the test phase was composed of 120 trials presented across 5 blocks of 24 trials.

### 2023-01-16 docs/manuscript.tex
continue responding to comments

@@
 
-% LD: if I talk about supervised/predictive learning at the end, I should probably include something here
 The most prominent theories of the inverse base-rate effect involve an attentional mechanism that drives not only learning but responding as well.
@@
 In terms of a clear observational-learning version of the IBRE, \citeA{johansen2007paradoxical} included the result of a short pilot experiment in their Appendix.
-There are no details about the procedure of this experiment, so we cannot make direct comparisons.
+Unfortunately, there is no statistical analysis confirming that $BC \to rare$ is reliably different than $BC \to common$.
+Given that sample size they report ($n = 16$), there is a chance that the study is underpowered.
+There are also no details about the procedure of this experiment.
+Therefore, we cannot make direct comparisons.
 Nonetheless, in their Experiment 2, \citeA{johansen2007paradoxical} tried to observe the inverse base-rate effect both in a predictive-learning condition and in an observational-learning condition.
@@
 Similarly, the $common$ category occured three times as often as $rare$.
-% LD: maybe include here some info on the shared-cue effect?
 Their design was optimized to investigate the hypothesised asymmetric cognitive representation of the two categories.
@@
 This provided evidence for the asymmetric cognitive representation \cite{kruschke2001inverse} hypothesised to develop during training.
-% LD: this speculation might be redundant?
 
-A simple neural-network explanation can account for \citeA{johansen2007paradoxical}'s.
-In a simple network, the rare feature develops stronger connections with the category label.
-Compound features share the connection with the category label, so any update to these connections dissipates between them.
-There is no need to relocate attention to reduce errors, therefore attention will not bias responding toward the rare label.
-But stronger $C \to rare$ feature-label connections will result in a rare bias on BC compounds during the test phase.
-This is strong evidence for the assymetric cognitive representation, but it still builds on the presence of explicit prediction error.
-In order to investigate the role of prediction error in response to feedback, we need to tweak their observational learning condition to conform to a more canonical implementation of the procedure.
-This will involve reimplementing the shared feature.
-Our current study modified the canonical IBRE procedure in an attempt to remove components from the design that is associated with explicit prediction error.
-This contrasts Experiment 2 in \citeA{johansen2007paradoxical}, who tried to remove the shared cue that pushed attention towards C during training in response to prediction error.
-Here, we strictly focus on prediction error as the element of the experimental procedure.
-% LD: I will also need to include Experiment 3 listed shared cue condition?
+Nonetheless, \citeA{johansen2007paradoxical} demonstrated that the inverse base-rate effect can occur without the traditional predictive learning design.
+In on of the conditions in their Experiment 3, the canonical inverse base-rate design (including the shared cue) was implemented in a list format.
+In this format, the trial-by-trial presentation of training items was turned into a list of 12 items fitted on a single page.
+In this condition, participants still exhibited the rare preference on $BC$ trials.
+In another condition of Experiment 3, participants recieved the information about outcome frequencies as a summary before testing.
+After learning about feature-label information in this manner, participants did not show the IBRE but was matching the base-rate.
+These experiments gives us evidence about another boundary condition for the IBRE - sequential presentation of training items.
+% Need to elaborate why this is relevant.
 
-% LD: I need to go through some studies because I think there is at least one more
-The only two studies directly looking at error-driven processes in the IBRE are \citeA{inkster2022neural} and \citeA{wills2014attention}.
+Three other studies directly looking at error-driven processes in the IBRE are \citeA{kruschke2005eye}, \citeA{inkster2022neural} and \citeA{wills2014attention}.
+\citeA{kruschke2005eye} demonstrated that attentional is allocated in line with
 \citeA{wills2014attention} observed posterior selection negativity and concurrent frontal positivity for C relative to B, which gave evidence for an error-driven selective attentional learning process.
@@
 In this study, we tested a central assumption of the most prominent theories of the IBRE.
-This central assumption was prediction error.
+This central assumption was that the IBRE is caused by the presence of prediction error.
 
-In our first attempt, we implemented an observational learning version of the canonical IBRE procedure.
+In our first experiment, we implemented an observational learning version of the canonical IBRE procedure.
 This meant that features and category labels appeared on the screen at the same time.
@@
 The experiment included no feedback and required no responses from participants during training.
-From a theoretical perspective, there was no opportunity for making an explicit error, therefore the condition for prediction error.
-Regardless, we observed the inverse base-rate effect.
+From a theoretical perspective, there was no opportunity for making an explicit error..
+Nevertheless, we observed the inverse base-rate effect.
 One limitation of this approach was that there are assumed causal relationships between features (symptoms) and labels (diseases).
-These relationships can predispose participants to make feature-to-label predictions, which will result in prediction error and attentional relocation.
+These relationships might predispose participants to make feature-to-label predictions, which will result in prediction error and attentional relocation.
 
-In our second attempt, we further removed the causal relationship between features and labels by changing the stimuli and their presentation.
+In our second experiment, we further removed the causal relationship between features and labels by changing the stimuli and their presentation.
 Here, participants saw nothing but an arrangement of geometric shapes, where previous category labels were treated as features.
 There were no causal links between features and labels.
-When participants needed to complete incomplete arrangements of these shapes, they still exhibited a rare bias on $BC$ trials.
+When participants were asked to complete incomplete arrangements of these shapes, they still exhibited a rare bias on $BC$ trials.
 We still observed the IBRE.
@@
 Most prominent theories and their corresponding formal specification rely on relocating attention in response to prediction error.
-They are unable to process the current experiment because they are not designed to encode information presented without feedback.
+They are unable to accommodate the current experiments because they are not designed to encode information presented without feedback.
+
+% LD: so it seems that Johansen et al treated asymmetric representation as an experimental condition and not as a theoretical construct
+% LD: I think this is a problem because asymmetric representation is hypothesized to be an asymmetry in the way the two categories represented
 
 The two experiments suggest that the necessary conditions to observe the IBRE are fewer than previously established.
-% LD: so it seems that Johansen et al treated asymmetric representation as an experimental condition and not as a theoretical construct
-% LD: I think this is a problem because asymmetric representation is hypothesized to be an asymmetry in the way the two categories represented
-Compared to \citeA{johansen2007paradoxical}, we also do not consider asymmetric representation and base-rate neglect as necessary conditions.
-We consider them to result from the environmental conditions specified in the experiment required for the IBRE as opposed to being necessary/sufficient conditions for the IBRE.
-From our point of view, they are theoretical constructs.
-They are not part of the experimental design.
-
 In Experiment 2, the only remaining conditions are the two uniquely predictive features, an overlapping feature, sequential presentation and the base rate.
-\cite{johansen2007paradoxical} gave evidence that even the overlapping features are not necessary.
-In their design, common category exemplars were made up of two features, while rare exemplars only had a single feature - disjoint cues.
-This can mean that common exemplars need to be composed of at least two features, while the rare exemplar must have only one feature.
-If the rare exemplar has one uniquely predictive feature and another feature that is not shared with the other category, we do not observe the effect.
-This is the shared-cue effect \cite{kruschke2001inverse, wills2014attention}.
-One can hypothesise that the shared cue and the disjoint cue are sufficient conditions, but at least one of them is necessary.
-In addition, one can hypothesise that in the absence of a shared feature, common exemplars must contain more features than rare exemplars.
-
-% Asymmetric representation can arise in different ways.
 One hypothesised way asymmetric representation is manifested is the attentional tuning of cognitive representation of category exemplars.
 This is not necessarily absent in our experiments but is not directly tested.
-Our experiments do not give direct evidence against the role of attention in developing asymmetric representation.
-Nonetheless, it must not happen through an error-driven process as conceptualised in most prominent theories.
+Our experiments do not give direct evidence against the role of attention in developing asymmetric representation or in its contributions to the emergence of the IBRE.
+Nonetheless, it must not happen through an error-driven process as conceptualised in most prominent theories of the IBRE.
 To further investigate this, the cued-recall procedure could incorporate eye-tracking to measure dwell time and order of information encoding.
-Even if we observe more and longer fixations to C relative to B, attention can still be allocated asymmetrically to drive the distinguishability of categories.
-Explanations do not need to invoke an explicit error-driven process.
+If we observe more and longer fixations on $C$ relative to $B$ on BC trials, it can mean that attention might be allocated asymmetrically to drive the distinguishability of categories.
+In that case, explanations do not need to invoke an explicit error-driven process.
+
+maybe a secondary cause of why ibre occurs?
 
@@
 We made available the two experiments written in javascript, the analysis code, the raw data, and all other supplementary materials.
-Experiment 1 is shared via \href{https://osf.io/auwvt/?view_only=2dc8384074fa4bcf9f2e3937fdaee2b4}{OSF}, and \href{https://github.com/lenarddome/ply216-observational-ibre}{GitHub}.
-Experiment 2 is shared via \href{https://osf.io/2tmc4/?view_only=489ebc888ef84a6b9b904072cfbe74df}{OSF} and \href{https://github.com/lenarddome/ply222-non-causal-ibre}{Github}.
-The main repository that includes this manuscript and links to the materials for the two experiments can be found on \href{https://github.com/lenarddome/pu093-errorless-attentionless-ibre}{GitHub}.
+Experiment 1 is shared via \href{https://osf.io/auwvt/?view_only=2dc8384074fa4bcf9f2e3937fdaee2b4}{OSF}, and \href{TBA}{GitHub}.
+Experiment 2 is shared via \href{https://osf.io/2tmc4/?view_only=489ebc888ef84a6b9b904072cfbe74df}{OSF} and \href{TBA}{Github}.
+The main repository that includes this manuscript and links to the materials for the two experiments can be found on \href{TBA}{GitHub}.
+
+\section{Acknowledgement}
+
+The data for Experiment 1 was collected as part of a final-year undergraduate project. We thank for the help and contributions of Charotte Hubbard, Selena Lockett, and Zoe Buller.
 

### 2023-01-16 docs/manuscript.tex
work on related work section

@@
 
-Three other studies directly looking at error-driven processes in the IBRE are \citeA{kruschke2005eye}, \citeA{inkster2022neural} and \citeA{wills2014attention}.
-\citeA{kruschke2005eye} demonstrated that attentional is allocated in line with
+Additionally, there are at least three studies which look at error-driven processes in the IBRE.
+\citeA{don2019learned} demonstrated that on $AC$ trials, people fixated on $C$ longer than on $A$ both pre-responding during stimulus presentation and post-responding during feedback.
+This fixation bias increased with more training.
+They also observe greater fixation on $C$ on $AC$ trials, relative to $B$ on $BC$ trials.
+Interestingly, there was no difference between fixation time of $C$ and $B$ during $BC$ trials.
+% still consistents with EXIT
 \citeA{wills2014attention} observed posterior selection negativity and concurrent frontal positivity for C relative to B, which gave evidence for an error-driven selective attentional learning process.
+These studies gave evidence that attentional reallocation occurs in line with the mechanisms of EXIT.
 \citeA{inkster2022neural} carried out a direct investigation into brain regions underlying error-driven learning in the IBRE.
@@
 They showed that these areas exhibited greater activation during the test phase for C relative to B with a presence of a shared cue during training.
-Both \citeA{wills2014attention} and \citeA{inkster2022neural} gave strong evidence in support of error-driven attentional learning accounts of the IBRE.
+It is reasonable to suggests that attentional occurs in a standard supervised learning paradigm and it is driven by prediction error.
+These three studies, \citeA{kruschke2005eye}, \citeA{wills2014attention} and \citeA{inkster2022neural} gave strong evidence in support of error-driven attentional learning accounts of the IBRE.
 In our study, we will look for the effect while trying reduce the chance that cognitive processes calculate prediction error through modifying the experimental design.
 In contrast, they looked for the neural substrates of error-driven processes incoporated into explanations of the effect.
+In sum, these studies provide support for an attentional reallocation mechanism that is mediated by prediction error.
 % This presumes that prediction error is necessary for the inverse base-rate effect.
@@
 Similarly, the test phase was composed of 120 trials presented across 5 blocks of 24 trials.
+\\
 
@@
 
-The data for Experiment 1 was collected as part of a final-year undergraduate project. We thank for the help and contributions of Charotte Hubbard, Selena Lockett, and Zoe Buller.
+The data for Experiment 1 was collected as part of a final-year undergraduate project. We thank for the help and contributions of To Be Added.
 

### 2023-01-17 docs/manuscript.tex
work on discussion

@@
 
+% LD: so it seems that Johansen et al treated asymmetric representation as an experimental condition and not as a theoretical construct
+% LD: I think this is a problem because asymmetric representation is hypothesized to be an asymmetry in the way the two categories represented
+
+The two experiments together suggest that the necessary conditions to observe the IBRE are fewer than previously established.
+In Experiment 2, the only remaining conditions are the two uniquely predictive features, an overlapping feature, sequential presentation and the base rate.
+One hypothesised way asymmetric representation is manifested is the attentional tuning of cognitive representation of category exemplars.
+This is not necessarily absent in our experiments but is not directly tested.
+Our experiments do not give direct evidence against the role of attention in developing asymmetric representation or in its contributions to the emergence of the IBRE.
+Nonetheless, it must not happen through an error-driven process as conceptualised in the most prominent theories of the IBRE.
+To further investigate this, the cued-recall procedure could incorporate eye-tracking to measure dwell time and order of information encoding.
+EXIT-like theories can informally predict longer fixations on $C$ relative to $B$ during training, but it is unclear what mechanism underlies this attentiontal allocation without an explicit error-driven process.
+
 In both experiments, the IBRE occurred without any explicit detail in the experimental procedure that would result in prediction error.
@@
 They are unable to accommodate the current experiments because they are not designed to encode information presented without feedback.
+Our results suggest that there could be a secondary cause of the IBRE not captured by previous process models.
 
-% LD: so it seems that Johansen et al treated asymmetric representation as an experimental condition and not as a theoretical construct
-% LD: I think this is a problem because asymmetric representation is hypothesized to be an asymmetry in the way the two categories represented
+Alternative theories that do not rely on prediction error might be able to accommodate our results.
+A version of the dissimilarity-similarity general context model \cite<DGCM, >{stewart2007dissimilarity} modified by \cite{o2018model} has been proposed as an explanation of the IBRE.
+From the perspective of DGCM, the main processes behind the rare preference are a combination of attention and dissimilarity from exemplars (stored category representations).
+$BC \to rare$ arises due to the combination of two factors: the high salience of $C$ relative to $B$ and the impact of the dissimilarity of $BC$ to the common category exemplar on the decision process.
+Nonetheless, the model can accommodate these results only informally because it does not specify the mechanisms which encode information and produce the attentional values of each stimulus.
+DGCM is a model of the test phase.
+In that sense, the current experimental manipulations cannot be represented in the model.
+This is the same challenge we encountered with process models - the specifications of these theories are unable to incorporate the changes to the experimental procedure.
 
-The two experiments suggest that the necessary conditions to observe the IBRE are fewer than previously established.
-In Experiment 2, the only remaining conditions are the two uniquely predictive features, an overlapping feature, sequential presentation and the base rate.
-One hypothesised way asymmetric representation is manifested is the attentional tuning of cognitive representation of category exemplars.
-This is not necessarily absent in our experiments but is not directly tested.
-Our experiments do not give direct evidence against the role of attention in developing asymmetric representation or in its contributions to the emergence of the IBRE.
-Nonetheless, it must not happen through an error-driven process as conceptualised in most prominent theories of the IBRE.
-To further investigate this, the cued-recall procedure could incorporate eye-tracking to measure dwell time and order of information encoding.
-If we observe more and longer fixations on $C$ relative to $B$ on BC trials, it can mean that attention might be allocated asymmetrically to drive the distinguishability of categories.
-In that case, explanations do not need to invoke an explicit error-driven process.
-
-maybe a secondary cause of why ibre occurs?
+Another alternative explanation is an eliminative-inference model \cite<ELMO, >{juslin2001high}.
+This approach considers the $BC \to rare$ bias to result from rule-learning and post-training inferential processes during $BC$ trials.
+The process is most similar to \textit{strategic guessing} \cite{kruschke1995extensions}.
+Because of the dissimilarity of $BC$ to the most frequently applied rule, $B \to common$, participants use the most similar rule applicable to $BC$ from a "guessing set", $C \to rare$.
+This results in the $BC \to rare$ bias.
+Informally, ELMO could accommodate the observational learning paradigm, because of the presence of feature-to-label causality.
+Participants could extract the same rules hypothesised to underlie the rare bias in the standard procedure.
+Due to this presumed causality, ELMO could hypothesise that people encode rules about symptoms and diseases, which will similarly result in $BC \to rare$ during test.
+But it is unclear how it could deal with the cued-recall implementation, as there is no clear-cut feature-to-label relationship in the stimuli presentation that drives rule formation.
+% One possibility is to assume that rules are formed between features only, but ELMO is not designed to deal with between-feature relationships.
 
@@
 
-Across two experiments, we investigated whether the IBRE requires error-driven components of the experimental procedure.
-In Experiment 1, we conducted a successful conceputal replication of \citeA{johansen2007paradoxical}, which gave evidence for the IBRE being independent of supervised learning procedures.
+Across two experiments, we investigated whether the demonstration of the IBRE requires the "guess-and-correct" components of the standard experimental procedure.
+In Experiment 1, we conducted a successful conceptual replication of \citeA{johansen2007paradoxical}, which gave evidence for the IBRE being independent of supervised learning procedures.
 In addition, Experiment 2 further suggests that the IBRE generalises beyond simple predictive-learning \cite<e.g.
 >{medin1988problem,kruschke1996base,wills2014attention} and decision-making \cite{johansen2007paradoxical} paradigms.
-This further suggest prediction error in terms of explicit feedback is not a necessary condition.
+This further suggests prediction error in terms of explicit feedback is not a necessary condition.
 Theories of IBRE are inadequate to account for these findings, because of their inability to extend beyond supervised learning.

### 2023-01-18 docs/manuscript.tex
add power calculation and tranisitonal edits

@@
 People tend to optimally label uniquely predictive features, $B$ and $C$, with their respective common and rare labels when presented by themselves.
-Responses on the shared feature $A$ tend to show base-rate following.
+Responses on the shared feature $A$ tend to show the base-rate following.
 But when uniquely predictive features are paired, $BC$, people tend to respond with the rare category label.
@@
 These theories are formal models.
-They are: a neural network with an exemplar-mediated attention to distinctive input, EXIT \cite{kruschke2001toward}, a three-layer neural network with competitive attentional gating, and a four-layer neural network with an additional rapid attentional shift \cite{paskewitz2020dissecting}.
+They are: a neural network with exemplar-mediated attention to distinctive input, EXIT \cite{kruschke2001toward}, a three-layer neural network with competitive attentional gating, and a four-layer neural network with an additional rapid attentional shift \cite{paskewitz2020dissecting}.
 All these explanations rely on a process that relocates attention in response to prediction errors - they update attentional values according to gradient descent.
@@
 During learning, people learn to label the $AB$ compound first, but they are still learning to label the $AC$ compound.
-The presence of $A$ tend to push participants to generalise what they learned about $AB$, so they label $AC$ as common, which results in an error.
+The presence of $A$ tends to push participants to generalise what they learned about $AB$, so they label $AC$ as common, which results in an error.
 After making this error, attention relocates towards the uniquely predictive feature $C$ to reduce future errors.
 This results in $C$ acquiring higher attentional salience than $B$.
-When the ambiguous $BC$ compound is presented, this attentional allocation presists and thus $C$ will dominate responding.
+When the ambiguous $BC$ compound is presented, this attentional allocation persists and thus $C$ will dominate responding.
 This results in an irrational tendency to respond with the rare label.
@@
 Unfortunately, there is no statistical analysis confirming that $BC \to rare$ is reliably different than $BC \to common$.
-Given that sample size they report ($n = 16$), there is a chance that the study is underpowered.
+Given the sample size they report, the study is underpowered \footnote{\citeA{johansen2007paradoxical} reports a sample size of $n = 16$.
+If we assume an effect size of $d = 0.46$ observed in \citeA{inkster2022neural} and an $\alpha = 0.05$ with a non-directional alternative hypothesis, the experiment has $power = 0.24$.
+We used the method provided by the R package pwr \cite{champely2020pwr} to calculate power.}.
 There are also no details about the procedure of this experiment.
@@
 According to the theories of the IBRE, this shared feature facilitates attentional relocation.
-This attentional tuning in turn pushes responding toward the rare label.
+This attentional tuning in turn pushes responding towards the rare label.
 In contrast, \citeA{johansen2007paradoxical} trained participants on $AB \to common$ and $C \to rare$.
-Similarly, the $common$ category occured three times as often as $rare$.
+Similarly, the $common$ category occurred three times as often as $rare$.
 Their design was optimized to investigate the hypothesised asymmetric cognitive representation of the two categories.
-As a result, the only instance when they observed a rare bias was when the common feature, $B$ ,presented in compound during training was paired with a rare, $C$, feature presented by itself during training.
-This provided evidence for the asymmetric cognitive representation \cite{kruschke2001inverse} hypothesised to develop during training.
+As a result, the only instance when they observed a rare bias was when the common feature, $B$, presented in compound during training was paired with a rare, $C$, feature presented by itself during training.
+This was taken as evidence for the asymmetric cognitive representation \cite{kruschke2001inverse} hypothesised to develop during training.
 
 Nonetheless, \citeA{johansen2007paradoxical} demonstrated that the inverse base-rate effect can occur without the traditional predictive learning design.
-In on of the conditions in their Experiment 3, the canonical inverse base-rate design (including the shared cue) was implemented in a list format.
+In one of the conditions in their Experiment 3, the canonical inverse base-rate design (including the shared cue) was implemented in a list format.
 In this format, the trial-by-trial presentation of training items was turned into a list of 12 items fitted on a single page.
-In this condition, participants still exhibited the rare preference on $BC$ trials.
-In another condition of Experiment 3, participants recieved the information about outcome frequencies as a summary before testing.
-After learning about feature-label information in this manner, participants did not show the IBRE but was matching the base-rate.
-These experiments gives us evidence about another boundary condition for the IBRE - sequential presentation of training items.
+Subsequently, participants made judgements about new cases on a separate page.
+In this condition, participants still exhibited a rare preference on $BC$ trials.
+In another condition of Experiment 3, participants received the information about outcome frequencies as a summary before testing.
+After learning about feature-label information in this manner, participants did not show the IBRE but was matching the base rate.
+These experiments give us evidence about another boundary condition for the IBRE - sequential presentation of training items.
 % Need to elaborate why this is relevant.
@@
 They also observe greater fixation on $C$ on $AC$ trials, relative to $B$ on $BC$ trials.
-Interestingly, there was no difference between fixation time of $C$ and $B$ during $BC$ trials.
+Interestingly, there was no difference between the fixation time of $C$ and $B$ during $BC$ trials.
 % still consistents with EXIT
@@
 They showed that these areas exhibited greater activation during the test phase for C relative to B with a presence of a shared cue during training.
-It is reasonable to suggests that attentional occurs in a standard supervised learning paradigm and it is driven by prediction error.
+It is reasonable to suggest that attentional occurs in a standard supervised learning paradigm and it is driven by prediction error.
 These three studies, \citeA{kruschke2005eye}, \citeA{wills2014attention} and \citeA{inkster2022neural} gave strong evidence in support of error-driven attentional learning accounts of the IBRE.
-In our study, we will look for the effect while trying reduce the chance that cognitive processes calculate prediction error through modifying the experimental design.
-In contrast, they looked for the neural substrates of error-driven processes incoporated into explanations of the effect.
+In our study, we will look for the effect while trying to reduce the chance that cognitive processes calculate prediction error by modifying the experimental design.
+In contrast, they looked for the neural substrates of error-driven processes incorporated into explanations of the effect.
 In sum, these studies provide support for an attentional reallocation mechanism that is mediated by prediction error.
@@
 Names were randomly drawn from a pool of male and female first names.
-The list was compiled from an online repository of popular baby names\footnote[1]{The list was taken and later curated from a GitHub repository: \href{https://github.com/aruljohn/popular-baby-names}{https://github.com/aruljohn/popular-baby-names}.}.
+The list was compiled from an online repository of popular baby names\footnote{The list was taken and later curated from a GitHub repository: \href{https://github.com/aruljohn/popular-baby-names}{https://github.com/aruljohn/popular-baby-names}.}.
 We selected the 50 most popular male and female names from 2021.

### 2023-01-18 docs/manuscript.tex
work on presenting power calculation

@@
 Unfortunately, there is no statistical analysis confirming that $BC \to rare$ is reliably different than $BC \to common$.
-Given the sample size they report, the study is underpowered \footnote{\citeA{johansen2007paradoxical} reports a sample size of $n = 16$.
-If we assume an effect size of $d = 0.46$ observed in \citeA{inkster2022neural} and an $\alpha = 0.05$ with a non-directional alternative hypothesis, the experiment has $power = 0.24$.
-We used the method provided by the R package pwr \cite{champely2020pwr} to calculate power.}.
+\citeA{johansen2007paradoxical} report a sample size of $n = 16$.
+If we use an effect size of $d = 0.46$ observed by \citeA{inkster2022neural} and an $\alpha = 0.05$ with a non-directional alternative hypothesis, the experiment has $power = 0.24$\footnote{We used the method provided by the R package pwr \cite{champely2020pwr} to calculate power.}.
+Given this information, the study is underpowered.
 There are also no details about the procedure of this experiment.
 Therefore, we cannot make direct comparisons.
+
 Nonetheless, in their Experiment 2, \citeA{johansen2007paradoxical} tried to observe the inverse base-rate effect both in a predictive-learning condition and in an observational-learning condition.
@@
 
-Additionally, there are at least three studies which look at error-driven processes in the IBRE.
+Additionally, there are at least three studies which directly look at error-driven processes in the IBRE.
 \citeA{don2019learned} demonstrated that on $AC$ trials, people fixated on $C$ longer than on $A$ both pre-responding during stimulus presentation and post-responding during feedback.
@@
 They also observe greater fixation on $C$ on $AC$ trials, relative to $B$ on $BC$ trials.
-Interestingly, there was no difference between the fixation time of $C$ and $B$ during $BC$ trials.
-% still consistents with EXIT
-\citeA{wills2014attention} observed posterior selection negativity and concurrent frontal positivity for C relative to B, which gave evidence for an error-driven selective attentional learning process.
-These studies gave evidence that attentional reallocation occurs in line with the mechanisms of EXIT.
+Furthermore, \citeA{wills2014attention} observed posterior selection negativity and concurrent frontal positivity for C relative to B, which gave evidence for an error-driven selective attentional learning process.
+These studies gave evidence that attentional reallocation occurs in line with the mechanisms of EXIT-like models.
 \citeA{inkster2022neural} carried out a direct investigation into brain regions underlying error-driven learning in the IBRE.
@@
 They showed that these areas exhibited greater activation during the test phase for C relative to B with a presence of a shared cue during training.
-It is reasonable to suggest that attentional occurs in a standard supervised learning paradigm and it is driven by prediction error.
-These three studies, \citeA{kruschke2005eye}, \citeA{wills2014attention} and \citeA{inkster2022neural} gave strong evidence in support of error-driven attentional learning accounts of the IBRE.
+Given these findings, it is reasonable to suggest that attentional reallocation occurs in a standard supervised learning paradigm and it is driven by prediction error.
+These three studies, \citeA{kruschke2005eye}, \citeA{wills2014attention} and \citeA{inkster2022neural} gave strong evidence in support of the error-driven attentional learning accounts of the IBRE.
+In sum, these studies provide support for an attentional reallocation mechanism that is mediated by prediction error.
 In our study, we will look for the effect while trying to reduce the chance that cognitive processes calculate prediction error by modifying the experimental design.
 In contrast, they looked for the neural substrates of error-driven processes incorporated into explanations of the effect.
-In sum, these studies provide support for an attentional reallocation mechanism that is mediated by prediction error.
 % This presumes that prediction error is necessary for the inverse base-rate effect.
@@
 Experiment 1 is a conceptual replication of an experiment included in the Appendix of \citeA{johansen2007paradoxical}.
-The only information available is the list of test items (23), the doubled-up design (2 sets of categories and features), and the sample size.
+The only information available from their report is the list of test items (23), the doubled-up design (2 sets of categories and features), and the sample size.
 We substantially simplified our implementation by removing the doubled-up design and reducing the number of test items to 6.
@@
 One limitation of this approach was that there are assumed causal relationships between features (symptoms) and labels (diseases).
-These relationships might predispose participants to make feature-to-label predictions, which will result in prediction error and attentional relocation.
+These relationships might predispose participants to make feature-to-label predictions, which will result in prediction error and attentional realllocation.
 

### 2023-01-18 docs/manuscript.tex
upgrade stats for experiment 2

@@
 The results are a qualitative and ordinal match to Experiment 1.
-Participants showed a clear common preference for stimuli A, $M_{A} = 0.77$, 95\% HDI $[0.68, 0.87]$, $\mathrm{BF}_{10} = 10,316.41$.
+Participants showed a clear common preference for stimuli A, $M = 0.74$, 95\% HDI $[0.66, 0.82]$, $\mathrm{BF}_{10} = 61,428.74$.
 
@@
 
-Participants also showed a reliable rare preference on ambiguous BC trials, $M_{BC} = 0.64$, 95\% HDI $[0.53, 0.75]$, $\mathrm{BF}_{10} = 4.09$
+Participants also showed a reliable rare preference on ambiguous BC trials, $M = 0.68$, 95\% HDI $[0.60, 0.77]$, $\mathrm{BF}_{10} = 412.77$.
 This gives us a sufficient amount of evidence to conclude that we have observed the IBRE.
@@
 The experiment included no feedback and required no responses from participants during training.
-From a theoretical perspective, there was no opportunity for making an explicit error..
+From a theoretical perspective, there was no opportunity for making an explicit error.
 Nevertheless, we observed the inverse base-rate effect.

### 2023-01-18 docs/manuscript.tex
managed to recover data from jatos after a server error - update stats

@@
 The results are a qualitative and ordinal match to Experiment 1.
-Participants showed a clear common preference for stimuli A, $M = 0.74$, 95\% HDI $[0.66, 0.82]$, $\mathrm{BF}_{10} = 61,428.74$.
+Participants showed a clear common preference for stimuli A, $M = 0.78$, 95\% HDI $[0.71, 0.85]$, $\mathrm{BF}_{10} = 7.72 \times 10^{8}$.
 
@@
 
-Participants also showed a reliable rare preference on ambiguous BC trials, $M = 0.68$, 95\% HDI $[0.60, 0.77]$, $\mathrm{BF}_{10} = 412.77$.
+Participants also showed a reliable rare preference on ambiguous BC trials, $M = 0.73$, 95\% HDI $[0.66, 0.80]$, $\mathrm{BF}_{10} = 581,370.03$.
 This gives us a sufficient amount of evidence to conclude that we have observed the IBRE.
@@
 In this study, we tested a central assumption of the most prominent theories of the IBRE.
-This central assumption was that the IBRE is caused by the presence of prediction error.
+This central assumption was that the IBRE is partially caused by the presence of prediction error.
 
@@
 One limitation of this approach was that there are assumed causal relationships between features (symptoms) and labels (diseases).
-These relationships might predispose participants to make feature-to-label predictions, which will result in prediction error and attentional realllocation.
+These relationships might predispose participants to make feature-to-label predictions, which will result in prediction error and attentional reallocation.
 
@@
 EXIT-like theories can informally predict longer fixations on $C$ relative to $B$ during training, but it is unclear what mechanism underlies this attentiontal allocation without an explicit error-driven process.
+In addition, brain imaging could further elaborate on the overlap of activations between cued-recall and supervised learning procedures.
+This would enable us to pinpoint the networks that uniquely underlie this rare preference but are independent of task demands.
 
-In both experiments, the IBRE occurred without any explicit detail in the experimental procedure that would result in prediction error.
+In both our experiments, the IBRE occurred without any explicit detail in the experimental procedure that would result in prediction error.
 Therefore, any theorised error-driven process must be able to operate without explicit feedback.
@@
 
-Alternative theories that do not rely on prediction error might be able to accommodate our results.
+There are alternative theories of the IBRE that do not rely on prediction-error driven processes.
 A version of the dissimilarity-similarity general context model \cite<DGCM, >{stewart2007dissimilarity} modified by \cite{o2018model} has been proposed as an explanation of the IBRE.

### 2023-01-18 docs/manuscript.tex
transitional commit before sending it off
correct sample size for experiment 2

@@
 
-We recruited 65 undergraduate students who completed the experiment for partial course credit.
+We recruited 122 undergraduate students who completed the experiment for partial course credit.
 Recruitment was done via the SONA recruitment system. \\
@@
 
-After exclusion, 30 participants made it into our analysis.
+After exclusion, 59 participants made it into our analysis.
 The group-level mean probabilities are shown in Table \ref{tab:results-exp2}.
@@
 
-There are alternative theories of the IBRE that do not rely on prediction-error driven processes.
+\subsection{Alternative Theories without Prediction Error}
+
+There are alternative theories of the IBRE that do not rely on processes that calculate prediction error.
 A version of the dissimilarity-similarity general context model \cite<DGCM, >{stewart2007dissimilarity} modified by \cite{o2018model} has been proposed as an explanation of the IBRE.

### 2023-01-25 docs/manuscript.tex
forgot to include an additional exclusion criteria for experiment 2

@@
 We applied the same analysis and exclusion methods as in Experiment 1.
+In addition, we also excluded participants whose mean reaction time fell one standard deviation, $\sigma = 519.96$ ms, below the group mean, $\mu = 1341.57$ ms.
 
@@
 A version of the dissimilarity-similarity general context model \cite<DGCM, >{stewart2007dissimilarity} modified by \cite{o2018model} has been proposed as an explanation of the IBRE.
-From the perspective of DGCM, the main processes behind the rare preference are a combination of attention and dissimilarity from exemplars (stored category representations).
+From the perspective of DGCM, the main processes behind the rare preference are a combination of attention, memory strenght of exemplars and dissimilarity from exemplars (stored category representations).
 $BC \to rare$ arises due to the combination of two factors: the high salience of $C$ relative to $B$ and the impact of the dissimilarity of $BC$ to the common category exemplar on the decision process.

### 2023-01-25 docs/manuscript.tex
finalise analysis

@@
 Therefore, we cannot make direct comparisons.
-
-Nonetheless, in their Experiment 2, \citeA{johansen2007paradoxical} tried to observe the inverse base-rate effect both in a predictive-learning condition and in an observational-learning condition.
-In the observational learning condition, category labels and features appeared together at the same time on the computer screen.
-Their design involved disjoint cues where categories shared no features in common.
-The canonical IBRE design involves a shared feature during training \cite{kruschke2001inverse, wills2014attention}.
-According to the theories of the IBRE, this shared feature facilitates attentional relocation.
-This attentional tuning in turn pushes responding towards the rare label.
-In contrast, \citeA{johansen2007paradoxical} trained participants on $AB \to common$ and $C \to rare$.
-Similarly, the $common$ category occurred three times as often as $rare$.
-Their design was optimized to investigate the hypothesised asymmetric cognitive representation of the two categories.
-As a result, the only instance when they observed a rare bias was when the common feature, $B$, presented in compound during training was paired with a rare, $C$, feature presented by itself during training.
-This was taken as evidence for the asymmetric cognitive representation \cite{kruschke2001inverse} hypothesised to develop during training.
 
@@
 
-We recruited 122 undergraduate students who completed the experiment for partial course credit.
+We recruited 171 undergraduate students who completed the experiment for partial course credit.
 Recruitment was done via the SONA recruitment system. \\
@@
 We applied the same analysis and exclusion methods as in Experiment 1.
-In addition, we also excluded participants whose mean reaction time fell one standard deviation, $\sigma = 519.96$ ms, below the group mean, $\mu = 1341.57$ ms.
 
@@
 
-After exclusion, 59 participants made it into our analysis.
+After exclusion, 86 participants made it into our analysis.
 The group-level mean probabilities are shown in Table \ref{tab:results-exp2}.
 The results are a qualitative and ordinal match to Experiment 1.
-Participants showed a clear common preference for stimuli A, $M = 0.78$, 95\% HDI $[0.71, 0.85]$, $\mathrm{BF}_{10} = 7.72 \times 10^{8}$.
+Participants showed a clear common preference for stimuli A, $M_{A} = 0.78$, 95\% HDI $[0.73, 0.83]$, $\mathrm{BF}_{10} = 5.37 \times 10^{13}$.
 
@@
 
-Participants also showed a reliable rare preference on ambiguous BC trials, $M = 0.73$, 95\% HDI $[0.66, 0.80]$, $\mathrm{BF}_{10} = 581,370.03$.
+Participants also showed a reliable rare preference on ambiguous BC trials, $M_{BC} = 0.73$, 95\% HDI $[0.67, 0.79]$, $\mathrm{BF}_{10} = 8.12 \times 10^{8}$.
 This gives us a sufficient amount of evidence to conclude that we have observed the IBRE.

### 2023-01-25 docs/manuscript.tex
add edits in response to comments

@@
 During learning, these sets of features occur at different frequencies.
-The features under the common label usually occur three times as much as features under the rare label \cite{kruschke1996base}.
+The features under the common label usually occur three times as often as features under the rare label \cite{kruschke1996base}.
 Following training, people categorise features presented by themselves and in novel combinations.
 People tend to optimally label uniquely predictive features, $B$ and $C$, with their respective common and rare labels when presented by themselves.
-Responses on the shared feature $A$ tend to show the base-rate following.
+Responses on the shared feature $A$ tend to show the base-rate following, $A \to common$.
 But when uniquely predictive features are paired, $BC$, people tend to respond with the rare category label.
 According to Classical Probability Theory, the rational response is to categorise this ambiguous combination under the common label, because it is the most frequently occurring label.
-This rare bias on ambiguous combinations of BC has been observed across a different variety of experimental manipulations \cite{kalish2001inverse,don2017effects,don2021attention,inkster2022effect,wills2014attention}.
+This rare bias on ambiguous combinations of BC has been observed across a variety of experimental manipulations \cite{kalish2001inverse,don2017effects,don2021attention,inkster2022effect,wills2014attention}.
 For a more thorough introduction to this irrational bias, see a review by \citeA{don2021hearing}.
@@
 
-The most prominent theories of the inverse base-rate effect involve an attentional mechanism that drives not only learning but responding as well.
+The most prominent theories of the inverse base-rate effect involve an an attentional meachanism that drives both learning and responding.
 These theories are formal models.
@@
 
-To our knowledge, there is only one attempt to implement the IBRE procedure without explicit feedback.
+To our knowledge, there is only one attempt to implement the standard trial-by-trial IBRE procedure without explicit feedback.
 In terms of a clear observational-learning version of the IBRE, \citeA{johansen2007paradoxical} included the result of a short pilot experiment in their Appendix.
-Unfortunately, there is no statistical analysis confirming that $BC \to rare$ is reliably different than $BC \to common$.
-\citeA{johansen2007paradoxical} report a sample size of $n = 16$.
-If we use an effect size of $d = 0.46$ observed by \citeA{inkster2022neural} and an $\alpha = 0.05$ with a non-directional alternative hypothesis, the experiment has $power = 0.24$\footnote{We used the method provided by the R package pwr \cite{champely2020pwr} to calculate power.}.
+Unfortunately, there is no statistical analysis confirming that the IBRE is reliably observed.
+\citeA{johansen2007paradoxical} report a sample size of 16.
+If we use an effect size of $d = 0.46$ observed by \citeA{inkster2022neural} and an $\alpha$ of 0.05 with a non-directional alternative hypothesis, the experiment has 24\% power\footnote{We used the method provided by the R package pwr \cite{champely2020pwr} to calculate power.}.
 Given this information, the study is underpowered.
@@
 After learning about feature-label information in this manner, participants did not show the IBRE but was matching the base rate.
-These experiments give us evidence about another boundary condition for the IBRE - sequential presentation of training items.
+These experiments give us evidence about another boundary condition for the IBRE - itemised rather than summerised presentation of training items.
 % Need to elaborate why this is relevant.
@@
 Additionally, there are at least three studies which directly look at error-driven processes in the IBRE.
-\citeA{don2019learned} demonstrated that on $AC$ trials, people fixated on $C$ longer than on $A$ both pre-responding during stimulus presentation and post-responding during feedback.
+\citeA{don2019learned}, demonstrated that on $AC$ trials, people fixated on $C$ longer than on $A$ both pre-responding during stimulus presentation and post-responding during feedback.
 This fixation bias increased with more training.
-They also observe greater fixation on $C$ on $AC$ trials, relative to $B$ on $BC$ trials.
-Furthermore, \citeA{wills2014attention} observed posterior selection negativity and concurrent frontal positivity for C relative to B, which gave evidence for an error-driven selective attentional learning process.
+They also observe greater fixation on $C$ on $AC$ trials, relative to $B$ on $AB$ trials.
+Furthermore, \citeA{wills2014attention} in an EEG study observed posterior selection negativity and concurrent frontal positivity for C relative to B, which gave evidence for an error-driven selective attentional learning process.
 These studies gave evidence that attentional reallocation occurs in line with the mechanisms of EXIT-like models.
 \citeA{inkster2022neural} carried out a direct investigation into brain regions underlying error-driven learning in the IBRE.
-Their ROI analysis explicitly targeted areas that were hypothesised to be involved in the computation of prediction error.
-They showed that these areas exhibited greater activation during the test phase for C relative to B with a presence of a shared cue during training.
-Given these findings, it is reasonable to suggest that attentional reallocation occurs in a standard supervised learning paradigm and it is driven by prediction error.
-These three studies, \citeA{kruschke2005eye}, \citeA{wills2014attention} and \citeA{inkster2022neural} gave strong evidence in support of the error-driven attentional learning accounts of the IBRE.
-In sum, these studies provide support for an attentional reallocation mechanism that is mediated by prediction error.
-In our study, we will look for the effect while trying to reduce the chance that cognitive processes calculate prediction error by modifying the experimental design.
-In contrast, they looked for the neural substrates of error-driven processes incorporated into explanations of the effect.
-% This presumes that prediction error is necessary for the inverse base-rate effect.
-% Arguably, one can predict that if there is no clear explicit design component (feedback) promoting prediction error as conceptualised in the models, the IBRE will not occur.
+Their region of interest (ROI) analysis explicitly targeted areas that were hypothesised to be involved in the computation of prediction error.
+They showed that these areas exhibited greater activation during the test phase for $C$ relative to $B$ with a presence of a shared cue during training.
+Given these findings, it is reasonable to suggest that prediction-error-driven attentional reallocation occurs in a standard supervised learning paradigm and it is driven by prediction error.
 
@@
 
-Below, we detail our first attempt to test whether we could observe the rare response bias without an explicit error-driven psychological mechanism.
+Below, we detail our first attempt to test whether we could observe the rare response bias to $BC$ without an explicit error-driven psychological mechanism.
 The design component which is most likely to result in any error-driven tuning is feedback.
@@
 % LD: work out how models might predict rational responding?
-
-Experiment 1 is a conceptual replication of an experiment included in the Appendix of \citeA{johansen2007paradoxical}.
-The only information available from their report is the list of test items (23), the doubled-up design (2 sets of categories and features), and the sample size.
 We substantially simplified our implementation by removing the doubled-up design and reducing the number of test items to 6.
@@
 They needed to complete reading the description within 5 seconds.
-If the 5 seconds threshold has passed, a screen appeared with the message '\textit{Please respond faster!}'.
+If the 5 seconds threshold was passed, a screen appeared with the message '\textit{Please respond faster!}'.
 In each training block, participants encountered 6 common diseases (common category exemplars) and 2 rare diseases (rare category exemplars).
@@
 In the test phase, participants judged individual symptoms and novel combinations of old symptoms, see Table \ref*{tab:abstract-exp1}.
-Symptoms appeared in a sentence, such as '\textit{John has a fever.}', with a prompt asking participants to label what disease the person has, '\textit{Does the patient have disease Z or disease L?}'.
+Symptoms appeared in a sentence, such as '\textit{John has a fever.}', with a prompt asking participants to say what disease the person has, '\textit{Does the patient have disease Z or disease L?}'.
 Participants had to respond by pressing either Z or L on the keyboard.
@@
 
-In order to test the presence of the IBRE, we calculated a Bayes Factor for a one-sample design.
+In order to test for the presence of the IBRE, we calculated a Bayes Factor for a one-sample design.
 We calculate the probability of responding with the rare label on the critical BC test item, $P(rare|BC)$, for each participant.
@@
 If the mean probability of $P(rare|BC)$ is higher than 0.5, we conclude that we observed the IBRE.
-Values lower than 0.5 would indicate rational responding.
+Values lower than 0.5 would indicate base-rate following.
 We used the method implemented in the BayesFactor R package \cite{morey2022bayes}.
@@
 
-One aspect of the current design is that participants can still experience internally-generated prediction errors from feature to categories on a trial-by-trial basis.
+One aspect of the current design is that participants might still experience internally-generated prediction errors from feature to categories on a trial-by-trial basis.
 Given that the general assumption is that diseases cause symptoms, participants could likely assume a causal link between symptoms and diseases.
-This assumed causal relationship can encourage participants to make not an explicit (responding with the category label via the keyboard) but an implicit prediction.
+This assumed causal relationship can encourage participants to make not an explicit (responding with the category label via the keyboard) but a silent prediction.
 Informally, participants might think of a certain feature--label causal relationship while reading the sentences.
@@
 Participants could respond by pressing either X or Y on the keyboard.
-Similarly, the test phase was composed of 120 trials presented across 5 blocks of 24 trials.
+The test phase was composed of 120 trials presented across 5 blocks of 24 trials.
 \\
@@
        & $P(common)$ & $P(rare)$ \\
-      \hline
-      A & 0.78 & 0.22  \\
-      AB & 0.95 & 0.05 \\
-      AC & 0.09 & 0.91 \\
-      B & 0.92 & 0.07  \\
-      \textbf{BC} & \textbf{0.35} & \textbf{0.65} \\
-      C & 0.08 & 0.92 \\
+      A & 0.78 & 0.22\\
+      AB & 0.95 & 0.05\\
+      AC & 0.08 & 0.92\\
+      B & 0.92 & 0.08\\
+      \textbf{BC} & \textbf{0.27} & \textbf{0.73}\\
+      C & 0.07 & 0.93\\
     \end{tabular}
@@
 In this study, we tested a central assumption of the most prominent theories of the IBRE.
-This central assumption was that the IBRE is partially caused by the presence of prediction error.
+This central assumption was that the IBRE is caused by the presence of prediction error.
 
@@
 This meant that features and category labels appeared on the screen at the same time.
-Participants learned about categories by reading complete sentences that describe what symptoms people exhibit and what diseases they have.
+Participants learned about categories by reading complete sentences that described what symptoms people exhibited and what diseases they had.
 The experiment included no feedback and required no responses from participants during training.
@@
 One limitation of this approach was that there are assumed causal relationships between features (symptoms) and labels (diseases).
-These relationships might predispose participants to make feature-to-label predictions, which will result in prediction error and attentional reallocation.
+These relationships might predispose participants to make feature-to-label predictions, which could result in prediction error and attentional reallocation.
 
@@
 There are alternative theories of the IBRE that do not rely on processes that calculate prediction error.
-A version of the dissimilarity-similarity general context model \cite<DGCM, >{stewart2007dissimilarity} modified by \cite{o2018model} has been proposed as an explanation of the IBRE.
+A version of the dissimilarity-similarity generalised context model \citeA<DGCM, >{stewart2007dissimilarity} modified by \citeA{o2018model} has been proposed as an explanation of the IBRE.
 From the perspective of DGCM, the main processes behind the rare preference are a combination of attention, memory strenght of exemplars and dissimilarity from exemplars (stored category representations).
@@
 This further suggests prediction error in terms of explicit feedback is not a necessary condition.
-Theories of IBRE are inadequate to account for these findings, because of their inability to extend beyond supervised learning.
+Theories of IBRE are inadequate to account for these findings, largely because of their inability to extend beyond supervised learning.
 
@@
 
-We made available the two experiments written in javascript, the analysis code, the raw data, and all other supplementary materials.
+We have made available the two experiments written in javascript, the analysis code, the raw data, and all other supplementary materials.
 Experiment 1 is shared via \href{https://osf.io/auwvt/?view_only=2dc8384074fa4bcf9f2e3937fdaee2b4}{OSF}, and \href{TBA}{GitHub}.
@@
 The main repository that includes this manuscript and links to the materials for the two experiments can be found on \href{TBA}{GitHub}.
+Links will be inserted when document is deanonymized.
 
@@
 
-The data for Experiment 1 was collected as part of a final-year undergraduate project. We thank for the help and contributions of To Be Added.
+The data for Experiment 1 was collected as part of a final-year undergraduate project. We thank for the help and contributions of [to be added on deanonymization of paper].
 

### 2023-01-25 docs/manuscript.tex
continue edits

@@
 In another condition of Experiment 3, participants received the information about outcome frequencies as a summary before testing.
+This summary was presented as prose.
 After learning about feature-label information in this manner, participants did not show the IBRE but was matching the base rate.
@@
 Additionally, there are at least three studies which directly look at error-driven processes in the IBRE.
-\citeA{don2019learned}, demonstrated that on $AC$ trials, people fixated on $C$ longer than on $A$ both pre-responding during stimulus presentation and post-responding during feedback.
+\citeA{don2019learned}, demonstrated that on $AC$ trials, people fixated on $C$ longer than on $A$ both pre-responding during stimulus presentation and post-responding during feedback (see also \citeA{kruschke2005eye}).
 This fixation bias increased with more training.
@@
 We used the method implemented in the BayesFactor R package \cite{morey2022bayes}.
+\\
 
@@
        & $P(common)$ & $P(rare)$ \\
+       \hline
       A & 0.78 & 0.22\\
@@
 
-% LD: so it seems that Johansen et al treated asymmetric representation as an experimental condition and not as a theoretical construct
-% LD: I think this is a problem because asymmetric representation is hypothesized to be an asymmetry in the way the two categories represented
-
 The two experiments together suggest that the necessary conditions to observe the IBRE are fewer than previously established.
@@
 
+% TODO: mention that ELMO can't do the shared cue effect
 Another alternative explanation is an eliminative-inference model \cite<ELMO, >{juslin2001high}.
@@
 But it is unclear how it could deal with the cued-recall implementation, as there is no clear-cut feature-to-label relationship in the stimuli presentation that drives rule formation.
+In addition, ELMO also predicts the IBRE in the absence of a shared cue, even though humans do not show the IBRE under those conditions \cite{kruschke2001inverse}.
 % One possibility is to assume that rules are formed between features only, but ELMO is not designed to deal with between-feature relationships.

### 2023-01-27 docs/manuscript.tex
transitional commit

@@
 
-The most prominent theories of the inverse base-rate effect involve an an attentional meachanism that drives both learning and responding.
+The most prominent theories of the IBRE involve an an attentional meachanism that drives both learning and responding.
 These theories are formal models.
@@
 Our overarching goal is to investigate whether we can still observe the IBRE, even if we experimentally remove a crucial assumption of already existing accounts.
-In our first attempt, building on the observational learning condition of Experiment 2 in \citeA{johansen2007paradoxical}, we implement the canonical IBRE design with a caveat that category labels are presented in unison with features.
+In our first attempt, we implement the canonical IBRE design with a caveat that category labels are presented in unison with features.
 
@@
 If we use an effect size of $d = 0.46$ observed by \citeA{inkster2022neural} and an $\alpha$ of 0.05 with a non-directional alternative hypothesis, the experiment has 24\% power\footnote{We used the method provided by the R package pwr \cite{champely2020pwr} to calculate power.}.
-Given this information, the study is underpowered.
+Given this information, this pilot experiment is underpowered.
 There are also no details about the procedure of this experiment.
@@
 Their region of interest (ROI) analysis explicitly targeted areas that were hypothesised to be involved in the computation of prediction error.
-They showed that these areas exhibited greater activation during the test phase for $C$ relative to $B$ with a presence of a shared cue during training.
+They showed that these areas exhibited greater activation during the test phase for $C$ relative to $B$.
 Given these findings, it is reasonable to suggest that prediction-error-driven attentional reallocation occurs in a standard supervised learning paradigm and it is driven by prediction error.
@@
 
+% FIXME: training trials ended after 10 seconds
 Table \ref{tab:abstract-exp1} summarises the abstract design of the experiment.
@@
 In the current experimental design, the IBRE emerged in the absence of an explicit prediction error that drives the development of attentional allocation.
-% This prediction error also drives the development of an asymmetric cognitive representation.
 All EXIT-like theories of the IBRE rely on the assumption that this irrational rare preference arises as a result of optimising accuracy during the training phase.
@@
 Given that the general assumption is that diseases cause symptoms, participants could likely assume a causal link between symptoms and diseases.
-This assumed causal relationship can encourage participants to make not an explicit (responding with the category label via the keyboard) but a silent prediction.
+This assumed causal relationship can encourage participants to make not an explicit but a silent prediction.
 Informally, participants might think of a certain feature--label causal relationship while reading the sentences.
@@
     \caption{Abstract design of Experiment 2 including both test and training phases.
-X and Y are in place of the category labels.
+X and Y are in place of the category labels common and rare.
 During the test phase, participants needed to select either X or Y to complete the features shown below.\\}
@@
 
-In the test phase, participants were shown \textit{incomplete} arrangement of geometric shapes and were asked to complete them.
+In the test phase, participants were shown \textit{incomplete} arrangements of geometric shapes and were asked to complete them.
 On each test trial, they were asked to select either \textbf{X} or \textbf{Y} to complete the arrangement.
@@
 A version of the dissimilarity-similarity generalised context model \citeA<DGCM, >{stewart2007dissimilarity} modified by \citeA{o2018model} has been proposed as an explanation of the IBRE.
-From the perspective of DGCM, the main processes behind the rare preference are a combination of attention, memory strenght of exemplars and dissimilarity from exemplars (stored category representations).
+From the perspective of DGCM, the main processes behind the rare preference are a combination of attention, memory strength of exemplars and dissimilarity from exemplars (stored category representations).
 $BC \to rare$ arises due to the combination of two factors: the high salience of $C$ relative to $B$ and the impact of the dissimilarity of $BC$ to the common category exemplar on the decision process.
@@
 
+% TODO: introduce guess-and-correct before the conclusion
 Across two experiments, we investigated whether the demonstration of the IBRE requires the "guess-and-correct" components of the standard experimental procedure.
@@
 The main repository that includes this manuscript and links to the materials for the two experiments can be found on \href{TBA}{GitHub}.
-Links will be inserted when document is deanonymized.
+GitHub links will be inserted when document is deanonymized.
 

### 2023-01-27 docs/manuscript.tex
move to US spelling
ther were no guidance about the spelling in the formatting requirement, but it seems everybody is using US spelling

@@
 But when uniquely predictive features are paired, $BC$, people tend to respond with the rare category label.
-According to Classical Probability Theory, the rational response is to categorise this ambiguous combination under the common label, because it is the most frequently occurring label.
+According to Classical Probability Theory, the rational response is to categorize this ambiguous combination under the common label, because it is the most frequently occurring label.
 This rare bias on ambiguous combinations of BC has been observed across a variety of experimental manipulations \cite{kalish2001inverse,don2017effects,don2021attention,inkster2022effect,wills2014attention}.
@@
 During learning, people learn to label the $AB$ compound first, but they are still learning to label the $AC$ compound.
-The presence of $A$ tends to push participants to generalise what they learned about $AB$, so they label $AC$ as common, which results in an error.
+The presence of $A$ tends to push participants to generalize what they learned about $AB$, so they label $AC$ as common, which results in an error.
 After making this error, attention relocates towards the uniquely predictive feature $C$ to reduce future errors.
@@
 According to these models, this irrationality results from an optimisation process that tries to reduce the errors people make.
-This process creates an asymmetric cognitive representation that can be summarised as $AB$ belongs to common, $AB \to common$, and C belongs to rare, $C \to rare$ \cite{kruschke2001inverse}.
+This process creates an asymmetric cognitive representation that can be summarized as $AB$ belongs to common, $AB \to common$, and C belongs to rare, $C \to rare$ \cite{kruschke2001inverse}.
 
@@
 After learning about feature-label information in this manner, participants did not show the IBRE but was matching the base rate.
-These experiments give us evidence about another boundary condition for the IBRE - itemised rather than summerised presentation of training items.
+These experiments give us evidence about another boundary condition for the IBRE - itemized rather than summerized presentation of training items.
 % Need to elaborate why this is relevant.
@@
 \citeA{inkster2022neural} carried out a direct investigation into brain regions underlying error-driven learning in the IBRE.
-Their region of interest (ROI) analysis explicitly targeted areas that were hypothesised to be involved in the computation of prediction error.
+Their region of interest (ROI) analysis explicitly targeted areas that were hypothesized to be involved in the computation of prediction error.
 They showed that these areas exhibited greater activation during the test phase for $C$ relative to $B$.
@@
 
-% FIXME: training trials ended after 10 seconds
-Table \ref{tab:abstract-exp1} summarises the abstract design of the experiment.
+Table \ref{tab:abstract-exp1} summarizes the abstract design of the experiment.
 This design is the simplest implementation of the IBRE procedure to date.
@@
 And features were selected to be solid black geometric shapes.
-The task asked participants to memorise the arrangement of these shapes.
-On each trial, we randomised the position of the geometric shapes in the arrangement.
-This further minimised the chances of having any design component suggestive of which feature is the category label.
+The task asked participants to memorize the arrangement of these shapes.
+On each trial, we randomized the position of the geometric shapes in the arrangement.
+This further minimized the chances of having any design component suggestive of which feature is the category label.
 
@@
 Similar to the previous experiment, participants completed two phases: an encoding/training and a test phase.
-In the training/encoding phase, participants were repeatedly exposed to the exemplars and were asked to memorise the arrangement of geometric shapes.
+In the training/encoding phase, participants were repeatedly exposed to the exemplars and were asked to memorize the arrangement of geometric shapes.
 Compared to Experiment 1, exemplars were composed of three geometric shapes.
@@
 In Experiment 2, the only remaining conditions are the two uniquely predictive features, an overlapping feature, sequential presentation and the base rate.
-One hypothesised way asymmetric representation is manifested is the attentional tuning of cognitive representation of category exemplars.
+One hypothesized way asymmetric representation is manifested is the attentional tuning of cognitive representation of category exemplars.
 This is not necessarily absent in our experiments but is not directly tested.
 Our experiments do not give direct evidence against the role of attention in developing asymmetric representation or in its contributions to the emergence of the IBRE.
-Nonetheless, it must not happen through an error-driven process as conceptualised in the most prominent theories of the IBRE.
+Nonetheless, it must not happen through an error-driven process as conceptualized in the most prominent theories of the IBRE.
 To further investigate this, the cued-recall procedure could incorporate eye-tracking to measure dwell time and order of information encoding.
@@
 In both our experiments, the IBRE occurred without any explicit detail in the experimental procedure that would result in prediction error.
-Therefore, any theorised error-driven process must be able to operate without explicit feedback.
+Therefore, any theorized error-driven process must be able to operate without explicit feedback.
 Most prominent theories and their corresponding formal specification rely on relocating attention in response to prediction error.
@@
 There are alternative theories of the IBRE that do not rely on processes that calculate prediction error.
-A version of the dissimilarity-similarity generalised context model \citeA<DGCM, >{stewart2007dissimilarity} modified by \citeA{o2018model} has been proposed as an explanation of the IBRE.
+A version of the dissimilarity-similarity generalized context model \citeA<DGCM, >{stewart2007dissimilarity} modified by \citeA{o2018model} has been proposed as an explanation of the IBRE.
 From the perspective of DGCM, the main processes behind the rare preference are a combination of attention, memory strength of exemplars and dissimilarity from exemplars (stored category representations).
@@
 
-% TODO: mention that ELMO can't do the shared cue effect
 Another alternative explanation is an eliminative-inference model \cite<ELMO, >{juslin2001high}.
@@
 
-% TODO: introduce guess-and-correct before the conclusion
 Across two experiments, we investigated whether the demonstration of the IBRE requires the "guess-and-correct" components of the standard experimental procedure.
 In Experiment 1, we conducted a successful conceptual replication of \citeA{johansen2007paradoxical}, which gave evidence for the IBRE being independent of supervised learning procedures.
-In addition, Experiment 2 further suggests that the IBRE generalises beyond simple predictive-learning \cite<e.g.
+In addition, Experiment 2 further suggests that the IBRE generalizes beyond simple predictive-learning \cite<e.g.
 >{medin1988problem,kruschke1996base,wills2014attention} and decision-making \cite{johansen2007paradoxical} paradigms.

### 2023-01-27 docs/manuscript.tex
transitional commit before final submission

@@
 From the perspective of DGCM, the main processes behind the rare preference are a combination of attention, memory strength of exemplars and dissimilarity from exemplars (stored category representations).
-$BC \to rare$ arises due to the combination of two factors: the high salience of $C$ relative to $B$ and the impact of the dissimilarity of $BC$ to the common category exemplar on the decision process.
+$BC \to rare$ arises due to the combination of the following factors: the high salience of $C$ relative to $B$ and the impact of the dissimilarity of $BC$ to the most remembered common category exemplar on the decision process.
 Nonetheless, the model can accommodate these results only informally because it does not specify the mechanisms which encode information and produce the attentional values of each stimulus.
@@
 
-Across two experiments, we investigated whether the demonstration of the IBRE requires the "guess-and-correct" components of the standard experimental procedure.
+Across two experiments, we investigated whether the demonstration of the IBRE requires the prediction-and-feedback components of the standard experimental procedure.
 In Experiment 1, we conducted a successful conceptual replication of \citeA{johansen2007paradoxical}, which gave evidence for the IBRE being independent of supervised learning procedures.

### 2023-04-17 docs/manuscript.tex
final copy

@@
 
-% \cogscifinalcopy % Uncomment this line for the final submission
+\cogscifinalcopy % Uncomment this line for the final submission
 
@@
 \author{{\large \bf Lenard Dome (lenarddome@gmail.com)} \\
-  Brain Research and Imaging Centre \\
-  University of Plymouth, Research Way, Plymouth, PL6 8BU
-  \AND {\large \bf Andy J. Wills (andy.wills@plymouth.ac.uk)} \\
-  Brain Research and Imaging Centre \\
+  {\large \bf Andy J. Wills (andy.wills@plymouth.ac.uk)} \\ \\
+  School of Psychology \\
   University of Plymouth, Research Way, Plymouth, PL6 8BU}
@@
 We have made available the two experiments written in javascript, the analysis code, the raw data, and all other supplementary materials.
-Experiment 1 is shared via \href{https://osf.io/auwvt/?view_only=2dc8384074fa4bcf9f2e3937fdaee2b4}{OSF}, and \href{TBA}{GitHub}.
-Experiment 2 is shared via \href{https://osf.io/2tmc4/?view_only=489ebc888ef84a6b9b904072cfbe74df}{OSF} and \href{TBA}{Github}.
-The main repository that includes this manuscript and links to the materials for the two experiments can be found on \href{TBA}{GitHub}.
-GitHub links will be inserted when document is deanonymized.
+Experiment 1 is shared via \href{https://osf.io/auwvt/}{https://osf.io/auwvt/}, and \href{https://github.com/lenarddome/ply216-observational-ibre}{https://github.com/lenarddome/ply216-observational-ibre}.
+Experiment 2 is shared via \href{https://osf.io/2tmc4/}{https://osf.io/2tmc4/} and \href{https://github.com/lenarddome/ply222-non-causal-ibreBA}{https://github.com/lenarddome/ply222-non-causal-ibre}.
+The main repository that includes this manuscript and links to the materials for the two experiments can be found on \href{https://github.com/lenarddome/pu093-errorless-attentionless-ibre}{https://github.com/lenarddome/pu093-errorless-attentionless-ibre}.
 
@@
 
-The data for Experiment 1 was collected as part of a final-year undergraduate project. We thank for the help and contributions of [to be added on deanonymization of paper].
+The data for Experiment 1 was collected as part of a final-year undergraduate project.
+We thank for the help and contributions of Selena Lockett and Charlotte Hubbard in the data collection for Experiment 1.
 

### 2023-04-17 docs/manuscript.tex
fix links

@@
 
-We have made available the two experiments written in javascript, the analysis code, the raw data, and all other supplementary materials.
-Experiment 1 is shared via \href{https://osf.io/auwvt/}{https://osf.io/auwvt/}, and \href{https://github.com/lenarddome/ply216-observational-ibre}{https://github.com/lenarddome/ply216-observational-ibre}.
-Experiment 2 is shared via \href{https://osf.io/2tmc4/}{https://osf.io/2tmc4/} and \href{https://github.com/lenarddome/ply222-non-causal-ibreBA}{https://github.com/lenarddome/ply222-non-causal-ibre}.
+We have made available the two experiments written in javascript, the analysis code, the raw data, and all other supplementary materials both on the Open Science Framework and GitHub.
+Experiment 1 is shared on \href{https://osf.io/auwvt/}{https://osf.io/auwvt/}, and \href{https://github.com/lenarddome/ply216-observational-ibre}{https://github.com/lenarddome/ply216-observational-ibre}.
+Experiment 2 is similarly shared on \href{https://osf.io/2tmc4/}{https://osf.io/2tmc4/} and \href{https://github.com/lenarddome/ply222-non-causal-ibreBA}{https://github.com/lenarddome/ply222-non-causal-ibre}.
 The main repository that includes this manuscript and links to the materials for the two experiments can be found on \href{https://github.com/lenarddome/pu093-errorless-attentionless-ibre}{https://github.com/lenarddome/pu093-errorless-attentionless-ibre}.

### 2023-04-17 docs/manuscript.tex
fix typo

@@
 \begin{abstract}
-
 The inverse base-rate effect is a robust irrational bias that arises when people face ambiguity.
 The most prominent theories of this irrational bias depend on prediction error.
-In this study, we gradually removed elements of a predictive learning design to test the extent to which error-driven processes underly this bias.
+In this study, we gradually removed elements of a predictive learning design to test the extent to which error-driven processes underlie this bias.
 In our first experiment, we removed explicit feedback by implementing the inverse base-rate effect in an observational learning procedure. 
